---
title: "Modelling - shark temperature and depth - trial model building"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("packages.R")
rm(list=ls()) # Clear out environment if necessary
```



```{r echo=FALSE, message=FALSE}
source(knitr::purl("../code/eda_shark_depth.Rmd", quiet=TRUE))
source(knitr::purl("../code/eda_shark_temp.Rmd", quiet=TRUE))
```


## To model temp vs depth, will need to do a few preliminary steps:

1. Decide temporal scale - daily, weekly, monthly?
2. Decide actual response - maximums, means, minimums? What about variability?
3. Wrangle data according to the above and ensure transmitter_id is kept in as we'll need it as a random effect in any model

## Need to bring in the individual T/P dataframe:
```{r}
source("individ_daily_td.R")
```


# Last few visual checks
```{r}
shark_1_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = daily_max_depth)) +
  geom_line(aes(x = date, y = daily_depth_sd), colour = "red") +
  facet_wrap( ~ transmitter_id)
```

## Check distributions so we know how to model
## Depth
```{r}
shark_1_td %>% 
  ggplot() +
  geom_histogram(aes(x = daily_max_depth))

shark_1_td %>% 
  ggplot() +
  geom_histogram(aes(x = daily_mean_depth))

shark_1_td %>% 
  ggplot() +
  geom_histogram(aes(x = daily_max_depth))

```

## Temp distribution
```{r}
shark_1_td %>% 
  ggplot() +
  geom_histogram(aes(x = daily_max_temp))
shark_1_td %>% 
  ggplot() +
  geom_histogram(aes(x = daily_mean_temp))
shark_1_td %>% 
  ggplot() +
  geom_histogram(aes(x = daily_max_temp))
```

```{r}
shark_1_td %>% 
  ggplot() +
  geom_histogram(aes(x = TL))
```


# Fit models

## Max depth ~ SST
```{r}
sst_maxd_glmm1 <- glmmTMB(daily_max_depth ~ SST + (1|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)
sst_maxd_glmm1a <- glmmTMB(daily_max_depth ~ SST + (SST|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)
sst_maxd_glmm2 <- glmmTMB(daily_max_depth ~ Season + (SST|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)
sst_maxd_glmm3 <- glmmTMB(daily_max_depth ~ SST * Season + (SST|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)
sst_maxd_glmm4 <- glmmTMB(daily_max_depth ~ SST * Season + Sex + (SST|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)
sst_maxd_glmm5 <- glmmTMB(daily_max_depth ~ SST * Season + Sex * TL + (SST|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)

```

## Residual diagnostics
```{r}
#plot_model(sst_maxd_glmm3,  type='diag')[-2] %>% plot_grid()
sst_maxd_glmm1 %>% simulateResiduals(plot=TRUE)
sst_maxd_glmm2 %>% simulateResiduals(plot=TRUE)
sst_maxd_glmm3 %>% simulateResiduals(plot=TRUE)
sst_maxd_glmm4 %>% simulateResiduals(plot=TRUE)
sst_maxd_glmm5 %>% simulateResiduals(plot=TRUE)
```
## Check AIC scores
```{r}
AIC(sst_maxd_glmm1, sst_maxd_glmm2, sst_maxd_glmm3, sst_maxd_glmm4, sst_maxd_glmm5)
```
## R2
```{r}
sst_maxd_glmm1 %>% performance::r2_nakagawa()
sst_maxd_glmm2 %>% performance::r2_nakagawa()
sst_maxd_glmm3 %>% performance::r2_nakagawa()
sst_maxd_glmm4 %>% performance::r2_nakagawa()
sst_maxd_glmm5 %>% performance::r2_nakagawa()
```
Model 5 best fit with fixed factors explaining 38% of the variance (marginal/pseudo R2) and lowest AIC. Also has all the required terms we are interested in. Will move forward with this model, but first note the random effect (individual - slope and intercept) is also explaining a chunk of the variance - indicates substantial individual level differences.

## Summary table
```{r}
summary(sst_maxd_glmm5)
```

# Summary of random effects
```{r ranfeffects}
sst_maxd_glmm5 %>% ranef()
```

# Quick overview of effects and interactions:

## Simple fixed effects
```{r}
sst_maxd_glmm5 %>% ggemmeans(~ Season) %>% plot
sst_maxd_glmm5 %>% ggemmeans(~ SST) %>% plot
sst_maxd_glmm5 %>% ggemmeans(~ Sex) %>% plot
sst_maxd_glmm5 %>% ggemmeans(~ TL) %>% plot
```
Looks like separately, there are some significant effects here:
- Effect of season: definitely seasonal differences - esp between winter and summer
- Effect of SST: a general increase with depth ~ SST
- Effect of sex: looks like some differences here but non-significant. Bayesian approach would probably allow quantification of this?
- Effect of size: again looks like there might be a pattern (smaller individuals going deeper), but this might be non-significant. Would need to do some planned contrasts to check.



## Interaction terms:

### Season * SST
```{r}
sst_maxd_glmm5 %>% ggemmeans(~ Season*SST) %>% plot
sst_maxd_glmm5 %>% ggemmeans(~ SST*Season) %>% plot
```
Quick interpretation - on average deeper in the summer, shallower in the winter and more variable in spring and autumn.

## Sex and size
```{r}
sst_maxd_glmm5 %>% ggemmeans(~ Sex*TL) %>% plot
sst_maxd_glmm5 %>% ggemmeans(~ TL*Sex) %>% plot
```

### SST and sex 
We know (from the above) that males may remain a little deeper than females, but do they respond differently to SSTs or in the same way?
```{r}
sst_maxd_glmm5 %>% ggemmeans(~ SST * Sex) %>% plot
sst_maxd_glmm5 %>% ggemmeans(~ Sex * SST) %>% plot
```

Looks like higher SSTs drive both males and females deeper

```{r}
sst_maxd_glmm3 %>% 
  emmeans( ~ Season) %>% 
  contrast() %>% 
  confint()
```


## What about just sex and size as a determinant of depths used?
Need Sex and Size to interact, with individual as random effect
```{r}
td_ss1 <- glmmTMB(daily_max_depth ~ Sex * TL + (1|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)
td_ss2 <- glmmTMB(daily_max_depth ~ Sex * TL + (SST|transmitter_id), data = shark_1_td, family = gaussian(), REML = TRUE)
```

## AIC
```{r}
AIC(td_ss1, td_ss2)
```


## Contrastss
```{r}
td_ss1 %>% ggemmeans(~ Sex) %>% plot
td_ss1 %>% ggemmeans(~ Sex * TL) %>% plot
td_ss2 %>% ggemmeans(~ Sex) %>% plot
td_ss2 %>% ggemmeans(~ Sex * TL) %>% plot
```
So it looks like overall, small males and females might avoid each other bit via depth, but overlap as they get older/bigger/to maturity. Makes sense? Need to conduct some planned contrasts I think if I want to state this..


