---
title: "Calibrating sensor data"
aauthor: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r packages, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(tidyverse)
```

# Housekeeping
```{r warning=FALSE}
rm(list=ls()) # Clear out environment if necessary
```

We start with the four separate files for teleosts and sharks, temperature and pressure as there are difference calibration values for each:


# Sharks #        

    Slope	    Intercept
P   0.9097	  -3.6388
T   0.1575	  -5.157

## Load shark data
```{r}
load(file = "../data/Rdata/shark_press_data.Rdata")
load(file = "../data/Rdata/shark_temp_data.Rdata")
```

## Calculate shark depth
```{r}
shark_press_data <- 
  shark_press_data %>% 
  mutate(Depth = `Sensor Value`*0.9097 - 3.6388)
```

## Calculate shark temperature
```{r}
shark_temp_data <- 
  shark_temp_data %>% 
  mutate(Temp = `Sensor Value`*0.1575 - 5.157)
```


# Combine together so that temp and depth in the same dataframes

## Combine sharks
```{r}
shark_tp_data <- 
  shark_temp_data %>% 
  full_join(shark_press_data) %>% 
  select(transmitter_id, detection_timestamp, receiver_name, Scientific_name, Date, 
         Location, Site, Sex, FL, TL, Lat, Long, Temp, Depth) %>% 
  arrange(detection_timestamp)
```

## Checks

### General
```{r}
shark_tp_data # 737000 observations
```

### How many individuals detected
```{r}
shark_tp_data %>% 
  filter(transmitter_id != 0) %>% 
  droplevels() %$% 
  summary(transmitter_id)
```
All 15 sharks with T/P tags show up at some point

### Check one or two individuals for both depth and temp
```{r}
shark_tp_data %>% 
  filter(transmitter_id == "A69-9004-14037") %>% 
  ggplot() +
  geom_point(aes(x = detection_timestamp, y = Depth)) +
  scale_y_reverse()

shark_tp_data %>% 
  filter(transmitter_id == "A69-9004-14037") %>% 
  ggplot() +
  geom_point(aes(x = detection_timestamp, y = Temp)) 
```

##################


# Teleosts #

     Slope	    Intercept
P    0.6065	  -2.4258
T    0.1575	  -5.157

## Load teleost data
```{r}
load(file = "../data/Rdata/teleost_press_data.Rdata")
load(file = "../data/Rdata/teleost_temp_data.Rdata")
```

## Calculate teleost depth
```{r}
teleost_press_data <- 
  teleost_press_data %>% 
  mutate(Depth = `Sensor Value`*0.6065 - 2.4258)
```

## Calculate teleost temperature
```{r}
teleost_temp_data <- 
  teleost_temp_data %>% 
  mutate(Temp = `Sensor Value`*0.1575 - 5.157)
```


## Combine teleosts
```{r}
teleost_tp_data <- 
  teleost_temp_data %>% 
  full_join(teleost_press_data) %>% 
  select(transmitter_id, detection_timestamp, receiver_name, Scientific_name, Date, 
         Location, Site, Sex, FL, TL, Lat, Long, Temp, Depth) %>% 
  arrange(detection_timestamp)
```




```{r}
save(shark_tp_data, file = "../data/Rdata/shark_tp_data.Rdata")
save(teleost_tp_data, file = "../data/Rdata/teleost_tp_data.Rdata")
```




