---
title: "Wrangling ROV data"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("packages.R")
#rm(list=ls()) # Clear out environment if necessary```
```

# Introduction

All ROV deployments in the Coral Sea have records of temperature and depth as part of deployment metadata log. 
We have deployments on Osprey from the summer of 2020/21 and 2022/23  -> sadly none from summer 2021/2022 which would correspond to our depth movement data
We also have access to temperature profile data from CTD drops in winter 2020 on RV Falkor - https://www.nodc.noaa.gov/archive/arc0163/0222330/1.1/data/0-data/FK200802_138578_ctd/data/FK200802_CTD010_20200820.hex
Need to process raw file using proprietary software - http://rstudio-pubs-static.s3.amazonaws.com/501186_8eb862a1b63242578e0b28468a4dadc9.html

## Load
### Deployment 22-02-2023 - Osprey North Horn - Dive 1 97m -> surface
```{r load-dep2023i, message=FALSE, warning=FALSE}
NHorn2023_D1 <- read_csv('../data/ROV/2023-02-22 08-35-42 vehicle1.csv') %>% 
  mutate(Time = as_hms(clock.currentTime)) %>% 
  select(Time, altitudeRelative, temperature.temperature1, temperature.temperature2) %>% 
  filter(Time > parse_hms("08:39:00")) %>%  # Beginning of the first 90m+ transect
  filter(Time < parse_hms("08:59:00"))  # End of dive
```

### Deployment 22-02-2023 - Osprey North Horn - Dive 2, 49m -> surface
```{r load-dep2023ii, message=FALSE, warning=FALSE}
NHorn2023_D2 <- read_csv('../data/ROV/2023-02-22 08-35-42 vehicle1.csv') %>% 
  mutate(Time = as_hms(clock.currentTime)) %>% 
  select(Time, altitudeRelative, temperature.temperature1, temperature.temperature2) %>% 
  filter(Time > parse_hms("09:11:30")) %>%  # ROV redeployed and sent back down to ~50m
  filter(Time < parse_hms("09:35:00"))  # End of dive
```

### Deployment 23-01-2021 - Osprey Entrance - Dive 1 95m - surface
```{r load-dep2021i, message=FALSE, warning=FALSE}
OspEnt2021 <- read_csv('../data/ROV/2021-02-23 11-32-50 vehicle1.csv') %>% 
  mutate(Time = as_hms(clock.currentTime)) %>% 
  select(Time, altitudeRelative, temperature.temperature1, temperature.temperature2)  %>% 
  filter(Time > parse_hms("12:02:30")) %>%  # Beginning of the first 95m+ transect
  filter(Time < parse_hms("13:20:30"))  # End of dive
```

### Deployment 24-01-2021 - Osprey North Horn - Dive 1 50m - surface
```{r load-dep2021ii, message=FALSE, warning=FALSE}
NHorn2021_D1 <- read_csv('../data/ROV/2021-02-24 07-43-14 vehicle1.csv') %>% 
  mutate(Time = as_hms(clock.currentTime)) %>% 
  select(Time, altitudeRelative, temperature.temperature1, temperature.temperature2)  %>% 
  filter(Time > parse_hms("09:08:30"))  %>%  # Beginning of the first ~50m transect
  filter(Time < parse_hms("09:40:30"))  # End of dive
```

### Deployment 24-01-2021 - Osprey North Horn - Dive 2 99 - 50m
```{r load-dep2021iii, message=FALSE, warning=FALSE}
NHorn2021_D2 <- read_csv('../data/ROV/2021-02-24 10-59-22 vehicle1.csv') %>% 
  mutate(Time = as_hms(clock.currentTime)) %>% 
  select(Time, altitudeRelative, temperature.temperature1, temperature.temperature2)  %>% 
  filter(Time > parse_hms("11:28:00"))  %>%  # Beginning of the first ~50m transect
  filter(Time < parse_hms("12:22:30"))  # End of dive
```


### Temp vs depth 

### Data from summer 2023 deployment
```{r}
#sumtemp2023 <- 
ggplot() +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2023_D1, colour = 'slateblue', size = 0.2) +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2023_D2, colour = 'navy', size = 0.2) +
  scale_x_continuous(limits = c(24, 31)) +
  theme_minimal() +
  xlab("Temp (째C)") + 
  ylab("Depth (m below sea level")

#sumtemp2023
```


### Data from summer 2021 deployment
```{r}
#sumtemp2021 <- 
ggplot() +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2021_D1, colour = 'grey70', size = 0.5) +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2021_D2, colour = 'grey50', size = 0.5) +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = OspEnt2021, colour = 'grey30', size = 0.5) +
  scale_x_continuous(limits = c(24, 31)) +
  theme_minimal() +
  xlab("Temp (째C)") + 
  ylab("Depth (m below sea level")

#sumtemp2021
```

# Combine all deployments and see if we can model a temp-depth relationship
```{r}
ROV_temp_comb <- 
  NHorn2021_D1 %>% 
  full_join(NHorn2021_D2) %>% 
  full_join(OspEnt2021) %>% 
  full_join(NHorn2023_D1) %>% 
  full_join(NHorn2023_D2) %>% 
  mutate(depth = altitudeRelative,
         temp = temperature.temperature2,
         .keep = "none") %>% 
  arrange(depth)
```

```{r}
ROV_temp_comb %>% 
  ggplot() +
  geom_point(aes(x = depth, y = temp), size = 0.2) +
  geom_smooth(aes(x = depth, y = temp))

ROV_temp_comb %>% 
  ggplot() +
  geom_point(aes(x = temp, y = depth), size = 0.2) +
  geom_smooth(aes(x = temp, y = depth),  method = "gam", formula = y ~ s(x, bs = 'cs'))

BLUEROVplot <- 
  ROV_temp_comb %>% 
  ggplot() +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2023_D1, colour = 'grey10', size = 0.3) +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2023_D2, colour = 'grey20', size = 0.3) +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2021_D1, colour = 'grey30', size = 0.3) +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = NHorn2021_D2, colour = 'grey45', size = 0.3) +
  geom_point(aes(x = temperature.temperature2, y = altitudeRelative), data = OspEnt2021, colour = 'grey60', size = 0.3) +
  #geom_smooth(aes(x = temp, y = depth)) +
  #geom_smooth(aes(x = temp, y = depth),  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.99, colour = 'red') +
  #geom_hline(yintercept = -50, colour = "blue", lty = '22') +
  #geom_vline(xintercept = 28.21, colour = "blue", lty = '22') +
  scale_x_continuous(limits = c(24.5, 30.5), breaks = c(24, 25, 26, 27, 28, 29, 30)) +
  theme_minimal() +
  xlab("Temp (째C)") + 
  ylab("Depth (m below sea level")

ROV_GAMplot <- 
  ROV_temp_comb %>% 
  ggplot() +
  #geom_smooth(aes(x = temp, y = depth)) +
  geom_smooth(aes(x = temp, y = depth),  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.99, colour = 'red') +
  geom_hline(yintercept = -50, colour = "blue", lty = '22') +
  geom_vline(xintercept = 28.21, colour = "blue", lty = '22') +
  scale_x_continuous(limits = c(24.5, 30.5), breaks = c(24, 25, 26, 27, 28, 29, 30)) +
  theme_minimal() +
  xlab("Temp (째C)") + 
  ylab("Depth (m below sea level")

BLUEROVplot + ROV_GAMplot
```



