---
title: "Load and wrangle satellite SST data"
aauthor: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 9
    fig_width: 16
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("code/packages.R")
#rm(list=ls()) # Clear out environment if necessary
getwd()
```

# Introduction

SST data for Osprey reef sourced from NOAA's Geostationary and Polar orbiting satellites

## Long term average monthly mean SSTs
### Dummy code in ymd for both 2021 and 2022, trimmed down in order to have as time series that matches dataset
```{r}
AMM_SST <- 
  read_excel('data/SST/osprey_climatology.xlsx', skip = 1) %>% 
  mutate(Month = as_date(paste(2021, Month, 15, sep = "-"))) %>%  # Use the middle/15th day of the month for plotting later
  bind_rows(read_excel('data/SST/osprey_climatology.xlsx', skip = 1) %>% 
              mutate(Month = as_date(paste(2022, Month, 15, sep = "-")))) %>% 
  filter(as_date(Month) >= as_date("2021-10-01"),
         as_date(Month) <= as_date("2022-10-31")) %>% 
  rename(AMM = SST)
```

## Plot to see long term average trends
```{r}
AMM_SST %>% 
  ggplot() +
  geom_line(aes(x = Month, y = AMM)) +
  geom_point(aes(x = Month, y = AMM)) +
  theme_minimal() +
  scale_x_date(date_breaks = "1 month", date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 30,  vjust = 0.8)) +
  xlab("Month of year") + 
  ylab("Averaged monthly mean SST \n (째C)")
```

## Load SST data for study period 
Study period - October 2021 - October 2022
```{r message = FALSE, warning = FALSE}
SST <- read_csv('data/SST/sst_geopolar_osprey_19850101-20230430.csv', skip = 5) %>% 
  mutate(date = ymd(dmy(Date)), .keep = "unused", .before = "SST") %>% 
  filter(#SST >0,
         as_date(date) >= as_date("2021-10-01"),
         as_date(date) <= as_date("2022-10-31"))
save(SST, file = "data/RData/SST.Rdata")
```


We obtained measurements for Osprey central (inside lagoon) as well as four other measurements c. 5km away from the reef in the cardinal directions, to check for any spatial differences in SST due to pooling/collection around reef. There were no differences of ecological importance:

```{r message=FALSE, warning=FALSE}
#SST_plot <- 
  ggplot() +
  geom_line(aes(x = date, y = SST), data = SST, colour = "black") +
  geom_line(aes(x = date, y = SST_N), data = SST, colour = "blue") +
  geom_line(aes(x = date, y = SST_S), data = SST, colour = "orange") +
  geom_line(aes(x = date, y = SST_E), data = SST, colour = "green4") +
  geom_line(aes(x = date, y = SST_W), data = SST, colour = "yellow") +
  theme_minimal() +
  scale_x_date(date_breaks = "1 month", date_labels = '%b-%y') +
  theme(axis.text.x = element_text(angle = 30,  vjust = 0.8)) +
  xlab("Month/year") + 
  ylab("Temp (째C)")
#SST_plot
#ggsave(filename = "output/SSTcomps.png", SST_plot)
```

## Plot study SST against AMM:
```{r}
ggplot() +
  geom_line(aes(x = Month, y = AMM), data = AMM_SST, col = "blue4") +
  geom_point(aes(x = Month, y = AMM), data = AMM_SST, col = "blue4", alpha =.5, size = 1) +
  geom_line(aes(x = date, y = SST), data = SST, colour = "black") +
  theme_bw() +
  scale_x_date(date_breaks = "1 month", date_labels = '%b-%y') +
  #theme(axis.text.x = element_blank()) +
  xlab("Month") + 
  ylab("Temp (째C)")
```

Would be good to plot the monthly mean for the study period against the AMM. Need to calculate first

```{r}
study_MM <- 
SST %>% 
  mutate(month = month(date, label = TRUE),
         year = year(date)) %>% 
  mutate(Month = as_date(paste(year, month, 15, sep = "-"))) %>%  # Use the middle/15th day of the month for plotting later
  group_by(year, Month) %>% 
  summarise(MM = mean(SST))%>% 
  left_join(AMM_SST) %>% 
  mutate(Month_diff = MM - AMM) 
study_MM
```


## Plot study SST against AMM again but now with MM for study period in
```{r  fig.width = 170/25.4, fig.height = 100/25.4}
SSTvAMMplot <- 
ggplot() +
  geom_line(aes(x = date, y = SST), data = SST, colour = "grey60") +
  geom_line(aes(x = Month, y = AMM), data = AMM_SST, colour = "dodgerblue") +
  geom_point(aes(x = Month, y = AMM), data = AMM_SST, fill = "dodgerblue", colour = "black",
             stroke = 0.8, pch = 22, alpha = 0.9, size = 2) +
  geom_line(aes(x = Month, y = MM), data = study_MM, colour = "firebrick") +
  geom_point(aes(x = Month, y = MM), data = study_MM, fill = "firebrick", colour = "black",
             stroke = 0.8, pch = 24, alpha = 0.9, size = 2) +
  theme_bw() +
  scale_x_date(date_breaks = "1 month", date_labels = '%b-%y', limits = c(as_date("2021-10-14"), as_date("2022-10-16"))) +
  theme(axis.text.x = element_text(angle = 90,  vjust = 3, hjust = 0.3)) +
  xlab("\nMonth/year") + 
  ylab("Temp (째C)\n")
SSTvAMMplot
```

# Save
```{r}
ggsave(filename = "output/SSTvAMMplot.pdf", SSTvAMMplot, width = 170, height = 100, units = "mm", dpi = 300)
```

## Hottest and coldest SSTs recorded for the study period
```{r}
SST %>% 
    filter(SST == max(SST))

SST %>% 
    filter(SST == min(SST))  
```



```{r}
# END
```


