---
title: "Wrangle and calibrate Osprey sensor data"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('code/packages.R')
#rm(list=ls()) # Clear out environment if necessary
getwd()
```

# Load data

```{r}
load("../CSMPWrangling/data/RData/alldata.RData")
```

## Filter out Osprey sensor data
```{r}
osp_sens_data <- 
  alldata %>% 
  filter(installation_name == 'Osprey') %>% 
  filter(Type != "pinger") %>% 
  droplevels()
```

# Checks
Should have 14 individual stations in the dataset
```{r}
summary(osp_sens_data$station_name_long)
```

## Check how many animals
Should have 25 tagged animals - compare database to project tag list.

### Load project tag list
```{r}
load("../CSMPWrangling/data/RData/project_tag_list.RData")
```

```{r}
osp_sens_data %>% 
  filter(Type == "temp")  %>% 
  group_by(transmitter_id, Scientific_name) %>% 
  distinct(Scientific_name)  %$% 
  summary(Scientific_name)

project_tag_list %>% 
  #filter(Type != "pinger") %$% 
  filter(Type == "temp")  %>% 
  droplevels() %$% 
  summary(Scientific_name)

```

# Missing 2 GTs
```{r}
project_tag_list %>% 
  filter(Org_type == "Teleost") %>% 
  filter(Type == "temp") %>% 
  #distinct(transmitter_id) %>% 
  anti_join(osp_sens_data) %>% 
  arrange(transmitter_id)
```

Two GTs tagged in March 2022 are completely missing from study. These were tagged in a slightly different location (Entrance - WPT120) so this might explain the absence..



## Check detection timing
```{r}
osp_sens_data %>% 
group_by(station_name_long) %>% 
  summarise(first_detection = min(detection_timestamp),
            last_detection = max(detection_timestamp)) %>% 
  mutate(deploy_duration = round(last_detection - first_detection, 0))
```


## Quick visual checks
```{r}
#osp_receiver_times <- 
osp_sens_data %>% 
  ggplot(mapping = aes(x = detection_timestamp, y = station_name_long)) + 
  xlab("Date") + 
  ylab("VR2W Location") +
  geom_point() +
  geom_vline(xintercept = as.numeric(osp_sens_data$detection_timestamp[c(5, 709325)])) +
  theme_light()
#osp_receiver_times
#ggsave(plot = osp_receiver_times, filename = "../output/osp_receiver_times.png", device = 'png', width = 180, height = 100, units = "mm")

```

Looks like we'll need to cut the data off at the end of Oct 2022 (black vertical line) as we only have 3 receivers with any data past that point

## Assign just one transmitter id to each sensor tag

Currently each animal has 2 x transmitter IDs for their tag - one for temp and one for pressure. If we want to to look at individual animal patterns we'll need to combine these. First will need to check which transmitter IDs go together, based on the serial number (which is the same for both T and P components):
```{r}
osp_sens_data %>% 
  dplyr:: select(transmitter_id, Serial, Org_type, Type) %>% 
  distinct() %>% 
  arrange(transmitter_id)
```

Sharks: a lower-even ID for temp goes with a higher-odd ID for press: Need to +1 to every temp/even ID
Teleosts: a lower-odd ID for temp goes with a higher-even ID for press: Need to +1 to every temp/odd ID

We'll do this and at the same time will calibrate using the calibration coefficients (slope and intercept) for the tags as provided by Vemco.

# Sharks
Calibration coefficients:

    Slope	    Intercept
P   0.9097	  -3.6388
T   0.1575	  -5.157

# Calculate temp
```{r}
shark_temp_data <- 
  osp_sens_data %>%
  filter(Org_type == "Elasmobranch") %>% 
  filter(ID %% 2 != 1)  %>% 
  #dplyr::select(transmitter_id) %>% 
  mutate(ID = ID+1) %>% 
  mutate(transmitter_id = factor(paste(Freq, Space, ID, sep = '-'))) %>% 
  mutate(Temp = `Sensor.Value` * Slope + Intercept)

```

# Calculate depth 
```{r}
shark_press_data <- 
  osp_sens_data %>%
  filter(Org_type == "Elasmobranch") %>% 
  filter(Type =="press") %>% 
  mutate(Depth = `Sensor.Value` * Slope + Intercept) %>% 
  mutate(Depth = if_else(Depth <0, 0, Depth)) # Have a few negative values (up to c. -3) - clearly cannot be out of the water so will reset to 0
```

# Combine together so that temp and depth in the same dataframes
```{r}
shark_tp_data <- 
  shark_temp_data %>% 
  full_join(shark_press_data) %>% 
  dplyr::select(transmitter_id, detection_timestamp, station_name, station_name_long, Scientific_name, Date_tagged, 
         Location_tagged, Site_tagged, Sex, FL, TL, Lat, Long, Temp, Depth) %>% 
  filter(!grepl("Lagoon", station_name_long),        # just a few shark detections (2 individs 14041 and 14059) from lagoon1 so filtering out
#         station_name_long != "Overnight_Mooring"   # may need to leave ONM out as it only has shallow/hot detections but will leave in for now
         ) %>%  
  filter(detection_timestamp < as.Date("2022-11-01 00:00:01")) %>% 
  arrange(detection_timestamp) %>% 
  droplevels()
```

## Add in other variables of interest for use in models later
- Hour of the day
- Time of day (day vs night)
- Season
```{r}
shark_tp_data <- shark_tp_data %>% 
  mutate(date = date(detection_timestamp)) %>% 
  mutate(hour = hour(detection_timestamp)) %>% 
  mutate(tod = case_when(hour %in% (7:16) ~ "Day",
                         hour %in% (5:6) ~ "Dawn",
                         hour %in% (17:18) ~ "Dusk",
                         TRUE ~ "Night")) %>% 
  mutate(Season = factor(case_when(
    date >= date("2022-01-01") & date<= date("2022-03-31") ~ "Summer",
    date >= date("2022-07-01") & date<= date("2022-09-30") ~ "Winter",
    date < date("2022-07-01") & date > date("2022-03-31") ~ "Autumn",
    date < date("2022-12-31")  ~ "Spring",
    date > date("2022-09-30") ~ "Spring"))) %>% 
  filter(date > date("2021-10-27")) %>% 
  mutate(Sex = factor(case_when(grepl("f", Sex) ~ "Female",
                                grepl("m", Sex) ~ "Male")))
```


# Checks
```{r}
shark_tp_data # No. of observations ~ 725,224
# Note this number is lower than the total # of shark observations as we've filtered out post-Oct2022
 osp_sens_data %>%
  filter(Org_type == "Elasmobranch") 
```


```{r}
shark_tp_data %$% summary(transmitter_id) # No. of individuals detected - all 15 sharks show up at some point & have 1000's of obs for each
#shark_tp_data  %>% distinct(transmitter_id) # Alternative summary
shark_tp_data %$% summary(station_name_long) # No. of individuals detected - all 15 sharks show up at some point & have 1000's of obs for each

```
# Save
```{r}
save(shark_tp_data, file = "data/Rdata/shark_tp_data.Rdata")
```


# Teleosts
Calibration coefficients:

     Slope	    Intercept
P    0.6065	  -2.4258ø
T    0.1575	  -5.157

# Calculate temp
```{r}
teleost_temp_data <- 
  osp_sens_data %>%
  filter(Org_type == "Teleost") %>% 
  filter(ID %% 2 == 1)  %>% 
  #dplyr::select(transmitter_id) %>% 
  mutate(ID = ID+1) %>% 
  mutate(transmitter_id = factor(paste(Freq, Space, ID, sep = '-'))) %>% 
  mutate(Temp = `Sensor.Value` * Slope + Intercept)
```

# Calculate depth
```{r}
teleost_press_data <- 
  osp_sens_data %>%
  filter(Org_type == "Teleost") %>% 
  filter(Type =="press") %>% 
  mutate(Depth = `Sensor.Value` * Slope + Intercept) %>% 
  mutate(Depth = if_else(Depth <0, 0, Depth)) # Have a few negative values (up to c. -3) - clearly cannot be out of the water so will reset to 0

```

# Combine together so that temp and depth in the same dataframes
```{r}
teleost_tp_data <- 
  teleost_temp_data %>% 
  full_join(teleost_press_data) %>% 
  select(transmitter_id, detection_timestamp, station_name, station_name_long, Scientific_name, Date_tagged, 
         Location_tagged, Site_tagged, Sex, FL, TL, Lat, Long, Temp, Depth) %>% 
  arrange(detection_timestamp) %>%   
  droplevels()
```

## Add in other variables of interest for use in models later
- Hour of the day
- Time of day (day vs night)
- Season
```{r}
teleost_tp_data <- teleost_tp_data %>% 
  mutate(date = date(detection_timestamp)) %>% 
  mutate(hour = hour(detection_timestamp)) %>% 
  mutate(tod = case_when(hour %in% (7:16) ~ "Day",
                         hour %in% (5:6) ~ "Dawn",
                         hour %in% (17:18) ~ "Dusk",
                         TRUE ~ "Night")) %>% 
  mutate(Season = factor(case_when(
    date >= date("2022-01-01") & date<= date("2022-03-31") ~ "Summer",
    date >= date("2022-07-01") & date<= date("2022-09-30") ~ "Winter",
    date < date("2022-07-01") & date > date("2022-03-31") ~ "Autumn",
    date < date("2022-12-31")  ~ "Spring",
    date > date("2022-09-30") ~ "Spring"))) %>% 
  filter(date > date("2021-10-27"))
```

# Checks
```{r}
teleost_tp_data # No. of observations ~ 41,000
teleost_tp_data %$% summary(transmitter_id) # No. of individuals detected - only 8 individuals. Already established we are missing 2 GTs
```

Only have good #s of detections for a couple of individuals.

# Save
```{r}
save(teleost_tp_data, file = "data/Rdata/teleost_tp_data.Rdata")
```


```{r}
#END
```

