---
title: "Exploratory data analysis - shark depth"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../packages.R")
#rm(list=ls()) # Clear out environment if necessary
```

# Load shark T/P data
```{r}
load(file = "../../data/Rdata/shark_tp_data.Rdata")
shark_tp_data
```

# Load SST data
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
setwd("../")
source("00_wrangle/wrangling_SSTdata.R")
```



## Depth overview - by individual
```{r fig.width=16, fig.height=6}
shark_tp_data %>% 
  ggplot() +
  geom_point(aes(x = detection_timestamp, y = Depth)) +
  scale_y_reverse() +
  facet_wrap(~ transmitter_id)
```

## Depth overview - by receiver station
```{r fig.width=16, fig.height=6}
shark_tp_data %>% 
  #filter(transmitter_id == "A69-9004-14037") %>%  # If you want to look at one individual
  ggplot() +
  geom_point(aes(x = detection_timestamp, y = Depth)) +
  scale_y_reverse() +
  facet_wrap(~ station_name_long)

shark_tp_data %>% 
  #filter(transmitter_id == "A69-9004-14037") %>%  # If you want to look at one individual
  ggplot() +
  geom_point(aes(x = detection_timestamp, y = Depth)) +
  scale_y_reverse() +
  facet_wrap(~ station_name)
```
Some receiver stations failed early in the deployment period

# Shark depth - daily summary
```{r}
shark_daily_depths <- 
  shark_tp_data %>% 
  mutate(date = date(detection_timestamp)) %>% 
  filter(!is.na(Depth)) %>% 
  group_by(date) %>% 
  summarise(daily_max_depth = max(Depth),
            daily_min_depth = min(Depth),
            daily_mean_depth = mean(Depth),
            daily_depth_sd = sd(Depth)) %>% 
  mutate(daily_min_depth = if_else(daily_min_depth <0, 0, daily_min_depth))
shark_daily_depths
#save(shark_daily_depths, file = "../data/Rdata/shark_daily_depths.Rdata")
```

## Plot daily depths
```{r fig.width=16, fig.height=6}
shark_daily_depth_plot <- 
shark_daily_depths %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  #geom_smooth(aes(x = date, y = daily_min_depth), colour = "blue",  method = "gam", formula = y ~ s(x, bs = 'cs')) + # method = "gam" implements mgcv
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_smooth(aes(x = date, y = daily_mean_depth), colour = "dodgerblue",  method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "red") + # Basic smooth, rather than gam
  theme_minimal() +
  scale_y_reverse() +
  #scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  #theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +
  #xlab("Date") +
  ylab("Depth \n (m below sealevel)")

shark_daily_depth_plot
```

# Weekly summary
```{r}
shark_weekly_depths <- 
shark_tp_data %>% 
  mutate(week_date = date(floor_date(detection_timestamp, "week"))) %>% 
  filter(!is.na(Depth)) %>% 
  group_by(week_date) %>% 
  summarise(weekly_max_depth = max(Depth),
            weekly_min_depth = min(Depth),
            weekly_mean_depth = mean(Depth),
            weekly_depth_sd = sd(Depth)) %>% 
  mutate(weekly_min_depth = if_else(weekly_min_depth <0, 0, weekly_min_depth))
shark_weekly_depths
```

## Plot weekly depths
```{r fig.width=16, fig.height=6}
shark_weekly_depth_plot <- 
shark_weekly_depths %>% 
  ggplot() +
  geom_line(aes(x = week_date, y = weekly_min_depth), colour = "blue") +
  geom_line(aes(x = week_date, y = weekly_mean_depth), colour = "grey30") +
  geom_smooth(aes(x = week_date, y = weekly_mean_depth), colour = "grey30", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  geom_line(aes(x = week_date, y = weekly_max_depth), colour = "red") +
  geom_smooth(aes(x = week_date, y = weekly_max_depth), colour = "red", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  scale_y_reverse() +
  theme_minimal() +
  xlab("Date") +
  ylab("Averaged weekly depths \n (m below sealevel)")
shark_weekly_depth_plot
```

# DEPTH PATTERNS BY INDIVIDUAL

## Individual daily depths
Going to add in SST data, Sex and TL at this point also as we'll need for modelling shortly
```{r}
shark_individual_daily_depths <- 
  shark_tp_data %>% 
  mutate(date = date(detection_timestamp)) %>% 
  filter(!is.na(Depth)) %>% 
  group_by(date, transmitter_id, Sex, TL) %>% 
  summarise(daily_max_depth = max(Depth),
            daily_min_depth = min(Depth),
            daily_mean_depth = mean(Depth),
            daily_depth_sd = sd(Depth))%>% 
  mutate(daily_min_depth = if_else(daily_min_depth <0, 0, daily_min_depth)) %>% 
  full_join(SST)
shark_individual_daily_depths
```

## Plot daily depths
```{r fig.width=16, fig.height=6}
shark_individual_daily_depth_plot <- 
shark_individual_daily_depths %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  #geom_smooth(aes(x = date, y = daily_min_depth), colour = "blue",  method = "gam", formula = y ~ s(x, bs = 'cs')) + # method = "gam" implements mgcv
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_smooth(aes(x = date, y = daily_mean_depth), colour = "dodgerblue",  method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "red") + # Basic smooth, rather than gam
  theme_minimal() +
  scale_y_reverse() +
  #scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  #theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +
  #xlab("Date") +
  ylab("Depth \n (m below sealevel)") +
  facet_wrap(facets = "transmitter_id")

shark_individual_daily_depth_plot
```


## Dau depths by sex
```{r fig.width=16, fig.height=6}
shark_individual_daily_depths %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  #geom_smooth(aes(x = date, y = daily_min_depth), colour = "blue",  method = "gam", formula = y ~ s(x, bs = 'cs')) + # method = "gam" implements mgcv
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_smooth(aes(x = date, y = daily_mean_depth), colour = "dodgerblue",  method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "red") + # Basic smooth, rather than gam
  theme_minimal() +
  scale_y_reverse() +
  #scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  #theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +
  #xlab("Date") +
  ylab("Depth \n (m below sealevel)") +
  facet_wrap(facets = "Sex")
```
Females much more variable in terms of depth use. Perhaps could be unbalanced sample design?

```{r}
shark_tp_data %>%
  select(transmitter_id, Sex) %>% 
  distinct() %$%
  summary(Sex)
```

Probably - 13 females and 2 males


# Weekly summary - by individual
```{r}
shark_individual_weekly_depths <- 
shark_tp_data %>% 
  mutate(week_date = date(floor_date(detection_timestamp, "week"))) %>% 
  filter(!is.na(Depth)) %>% 
  group_by(week_date, transmitter_id, Sex, TL) %>% 
  summarise(weekly_max_depth = max(Depth),
            weekly_min_depth = min(Depth),
            weekly_mean_depth = mean(Depth),
            weekly_depth_sd = sd(Depth))%>% 
  mutate(weekly_min_depth = if_else(weekly_min_depth <0, 0, weekly_min_depth))

shark_individual_weekly_depths
```

## Plot individual weekly depths
```{r fig.width=16, fig.height=6}
shark_individual_weekly_depth_plot <- 
shark_individual_weekly_depths %>% 
  ggplot() +
  geom_line(aes(x = week_date, y = weekly_min_depth), colour = "blue") +
  geom_line(aes(x = week_date, y = weekly_mean_depth), colour = "grey30") +
  geom_smooth(aes(x = week_date, y = weekly_mean_depth), colour = "grey30", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  geom_line(aes(x = week_date, y = weekly_max_depth), colour = "red") +
  geom_smooth(aes(x = week_date, y = weekly_max_depth), colour = "red", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  scale_y_reverse() +
  theme_minimal() +
  xlab("Date") +
  ylab("Averaged weekly depths \n (m below sealevel)") +
  facet_wrap(facets = "transmitter_id")

shark_individual_weekly_depth_plot
```




