---
title: "Shark data: daily"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_width: 161/25.4
    fig_height: 100/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("code/packages.R")
source("code/extras.R")
#rm(list=ls()) # Clear out environment if necessary
```

# Load data 
## Shark T/P data
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
load(file = "data/Rdata/shark_tp_data.Rdata")
# Or, if you also want all the associated files loaded:
#source(knitr::purl("code/00_wrangle/wrangling_sensordata.Rmd", quiet = TRUE))
```

## SST
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
load(file = "data/RData/SST.Rdata")
# Or, if you also want all the associated files loaded:
#source(knitr::purl("code/00_wrangle/wrangling_SSTdata.Rmd", quiet = TRUE))
```


# Shark depth - daily summary
```{r}
shark_daily_depths <- 
  shark_tp_data %>% 
  filter(!is.na(Depth)) %>% 
  group_by(date) %>% 
  summarise(daily_max_depth = max(Depth),
            daily_min_depth = min(Depth),
            daily_mean_depth = mean(Depth),
            daily_depth_sd = sd(Depth)) %>% 
  mutate(daily_min_depth = if_else(daily_min_depth <0, 0, daily_min_depth))
#shark_daily_depths
#save(shark_daily_depths, file = "data/Rdata/shark_daily_depths.Rdata")
```

# Shark temps - daily summary
```{r}
shark_daily_temps <- 
  shark_tp_data %>% 
  filter(!is.na(Temp)) %>% 
  group_by(date) %>% 
  summarise(daily_max_temp = max(Temp),
            daily_min_temp = min(Temp),
            daily_mean_temp = mean(Temp),
            daily_temp_sd = sd(Temp))
#shark_daily_temps
#save(shark_daily_temps, file = "data/Rdata/shark_daily_temps.Rdata")
```


# Combine
```{r}
shark_daily_td <- 
  SST %>% 
  full_join(shark_daily_depths) %>% 
  full_join(shark_daily_temps) %>% 
  filter(date > as.Date("2021-10-29")) %>% 
  mutate(daily_temp_diff = daily_max_temp - daily_min_temp,
         daily_depth_diff = daily_max_depth -daily_min_depth)
#save(shark_daily_td, file = "data/Rdata/shark_daily_td.Rdata")
```


# 1. OVERVIEW #

## SST only ##

```{r fig.width = 140/25.4, fig.height = 40/25.4}
sst_2021_2022 <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.6, lty = 3313) +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  xlab("\nMonth of year (2021 - 2022)") +
  ylab("SST (°C)") +
  theme_bjc() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank()) 

# View
sst_2021_2022

# Save
#ggsave(filename = "output/sst_2021_2022_fig.pdf", plot = sst_2021_2022,
#       device = cairo_pdf, width = 140, height = 40, units = "mm", dpi = 300)
```


## Daily temps (all) vs SST
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_temps_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.6, lty = 3313) +
  geom_line(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_smooth(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_line(aes(x = date, y = daily_min_temp), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_min_temp), colour = "midnightblue") +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  geom_smooth(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  theme_bjc() +
  theme(axis.text.x = element_text(angle = 0)) +
  xlab("\nMonth of year (2021 - 2022)") +
  ylab("Temp (°C)")

# View
shark_daily_temps_plot

# Save
#ggsave(filename = "output/shark_daily_temps_plot.pdf", plot = shark_daily_temps_plot, 
 #      device = cairo_pdf, width = 140, height = 70, units = "mm", dpi = 300)
```

# 2. DAILY MAXIMA #

# Daily maximumm temperatures track closely with SST
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_maxtemp_plot <- 
  shark_daily_td %>% 
  ggplot() +
  #geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.5) +
  geom_line(aes(x = date, y = daily_max_temp),  colour = "red") +
  #geom_smooth(aes(x = date, y = daily_max_temp), colour = "red") +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  scale_y_continuous(limits = c(23, 31), breaks = c(23, 24, 25, 26, 27, 28, 29, 30, 31)) +
  theme_bjc() +
  xlab("Date") + 
  ylab(expression(paste(italic("T"),""[b]* " (°C)"))) 

# View
shark_daily_maxtemp_plot
```


```{r fig.width = 140/25.4, fig.height = 70/25.4}
# Save
#ggsave(filename = "output/shark_daily_maxtemp_plot.pdf", plot = shark_daily_maxtemp_plot, 
 #      device = cairo_pdf, width = 140, height = 70, units = "mm", dpi = 300)
```

### Quick plot of relationship between SST and daily max temps (strong relationship)
```{r fig.width = 70/25.4, fig.height = 70/25.4}
shark_SST_maxtemp_lm_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_max_temp)) +
  geom_smooth(aes(x = SST, y = daily_max_temp), method = "glm") +
  theme_bjc() +
  ylab("Daily max temp (°C)") + 
  xlab("SST (°C)")

# View  
shark_SST_maxtemp_lm_plot

# Save - don't need to save as we model properly later
#ggsave(filename = "output/shark_SST_mintemp_lm_plot.pdf", plot = shark_SST_mintemp_lm_plot, width = 100, height = 100, units = "mm", dpi = 300)
```

# 3. DAILY MEANS #

# However, during summer months sharks were able to negatively thermoregulate mean body temperatures despite high SSTs:
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_meantemp_plot <- 
  shark_daily_td %>% 
  ggplot() +
  #geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.5) +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #geom_smooth(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #geom_vline(xintercept = as.Date("2022-01-19"), colour = "red", lwd = 1) +
 # geom_segment(aes(x = as.Date("2022-01-19"), xend = as.Date("2022-01-19"), y = 29.6, yend = 28.4), colour = "blue", arrow = arrow(length = unit(0.2, "cm"))) +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  scale_y_continuous(limits = c(23, 31), breaks = c(23, 24, 25, 26, 27, 28, 29, 30, 31)) +
  theme_bjc() +
  xlab("Date") + 
  ylab(expression(paste(italic("T"),""[b]* " (°C)"))) 

# View
shark_daily_meantemp_plot
```

```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_alltb_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = daily_max_temp),  colour = "red") +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  geom_line(aes(x = date, y = daily_min_temp),  colour = "blue") +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  scale_y_continuous(limits = c(23, 31), breaks = c(23, 24, 25, 26, 27, 28, 29, 30, 31)) +
  theme_bjc() +
  xlab("Date") + 
  ylab(expression(paste(italic("T"),""[b]* " (°C)"))) 

# shark_daily_alltb_plot
shark_daily_alltb_plot
```



```{r fig.width = 140/25.4, fig.height = 70/25.4}
# Save
#ggsave(filename = "output/shark_daily_alltb_plot.pdf", plot = shark_daily_alltb_plot, 
#       device = cairo_pdf, width = 140, height = 70, units = "mm", dpi = 300)
```

# The maximum daily downward thermo-regulation was as much as 2°C (on 19 Jan 2022), when SST was around 30°C. Max depth recorded on this date was 122m. (Potentially could investigate what this would come out as at individual level?)
```{r}
#shark_daily_td <- 
#  shark_daily_td %>% 
#  mutate(temp_diff = daily_max_temp - daily_mean_temp)

shark_daily_td %>% 
  summarise(max_temp_diff = max(daily_temp_diff))
  #filter(daily_temp_diff == max(temp_diff))
  
shark_daily_td %>% 
  slice_max(daily_temp_diff)
```

## When plotted over the course of the year, it is clear that these diffs are generally higher in the summer
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = daily_temp_diff)) +
  geom_smooth(aes(x = date, y = daily_temp_diff)) +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_bjc() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("\nDate") + 
  ylab("Daily temp diff\n(max vs mean daily temp, °C)")
```

### Quick plot of relationship between SST and daily mean temps (there is one - will have to check effect size in model later)
```{r fig.width = 70/25.4, fig.height = 70/25.4}
shark_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_mean_temp)) +
  geom_smooth(aes(x = SST, y = daily_mean_temp), method = "glm") +
  theme_bjc() +
  ylab("Daily mean temp (°C)") + 
  xlab("SST (°C)")
```

# Can also look at mean temp SD ~ SST
```{r fig.width = 70/25.4, fig.height = 70/25.4}
shark_daily_td %>% 
  ggplot() +
  #geom_point(aes(x = SST, y = temp_diff)) +
  #geom_smooth(aes(x = SST, y = temp_diff), method = "glm") +
  geom_point(aes(x = SST, y = daily_temp_sd)) +
  geom_smooth(aes(x = SST, y = daily_temp_sd), method = "gam") +
  theme_bjc() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("SST") + 
  ylab("Daily temp diff\n(max vs mean daily temp, °C)")
```



## Relationship between SST


## Minimum daily temperatures were consistent throughout the year and did not depend on SST (25.3°C ± 0.6°C SD)
### Calculate avg and SD
```{r}
shark_daily_td %>% 
  summarise(avg_daily_min_temp = mean(daily_min_temp),
            sd_daily_min_temp = sd(daily_min_temp))
```

### Plot relationship between SST and daily min temps (there isn't one)
```{r fig.width = 70/25.4, fig.height = 70/25.4}
shark_SST_mintemp_lm_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_min_temp)) +
  geom_smooth(aes(x = SST, y = daily_min_temp), method = "glm") +
  theme_bjc() +
  ylab("Daily min temp (°C)") + 
  xlab("SST (°C)")

# View  
shark_SST_mintemp_lm_plot

# Save - don't need to save as we model properly later
#ggsave(filename = "output/shark_SST_mintemp_lm_plot.pdf", plot = shark_SST_mintemp_lm_plot, width = 100, height = 100, units = "mm", dpi = 300)
```


```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_maxmintemp_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = daily_max_temp),  colour = "red") +
  geom_line(aes(x = date, y = daily_min_temp),  colour = "blue") +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  scale_y_continuous(limits = c(23, 31), breaks = c(23, 24, 25, 26, 27, 28, 29, 30, 31)) +
  theme_bjc() +
  xlab("Date") + 
  ylab(expression(paste(italic("T"),""[b]* " (°C)"))) 

# View
shark_daily_maxmintemp_plot
```


```{r fig.width = 140/25.4, fig.height = 70/25.4}
# Save
#ggsave(filename = "output/shark_daily_maxmintemp_plot.pdf", plot = shark_daily_maxmintemp_plot, 
  #     device = cairo_pdf, width = 140, height = 70, units = "mm", dpi = 300)

```


```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_mintemp_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = daily_min_temp),  colour = "blue") +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  scale_y_continuous(limits = c(23, 31), breaks = c(23, 24, 25, 26, 27, 28, 29, 30, 31)) +
  theme_bjc() +
  xlab("\nDate") + 
  ylab("Temp (°C)")

# View
shark_daily_mintemp_plot
```

# Combine figs
```{r}
daily_min_temp_sum_fig <- 
  ggarrange(shark_daily_mintemp_plot, shark_SST_mintemp_lm_plot, ncol = 2, widths = c(1.6,1))
```

# And save
```{r}
#ggsave(filename = "output/daily_min_temp_sum_fig.pdf", plot = daily_min_temp_sum_fig, 
#       width = 180, height = 60, units = "mm", dpi = 300)
```


## Downward thermoregulation appears to correspond with maximum daily depths attained - depth movements used in order to thermoregulate
```{r fig.width = 70/25.4, fig.height = 70/25.4}
shark_daily_td %>% 
  ggplot() +
  geom_point(aes(x = daily_temp_diff, y = daily_max_depth)) +
  #geom_smooth(aes(x = daily_max_temp, y = daily_max_depth), method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_smooth(aes(x = daily_temp_diff, y = daily_max_depth), method = "lm") +
  theme_bjc() +
  ylab("Daily maximum depth (mbsl)") + 
  xlab("Daily temperature difference (°C)")
```

... so let's explore depths a bit more


## Plot daily depths
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_depth_plot <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  #geom_smooth(aes(x = date, y = daily_min_depth), colour = "blue",  method = "gam", formula = y ~ s(x, bs = 'cs')) + # method = "gam" implements mgcv
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_smooth(aes(x = date, y = daily_mean_depth), colour = "dodgerblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "red") + # Basic smooth, rather than gam
  theme_bjc() +
  scale_y_reverse() +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  theme(axis.text.x = element_text(angle = 0)) +
  #theme(axis.text.x = element_blank()) +
  #theme(axis.title.x = element_blank()) +
  #theme(axis.ticks.x = element_blank()) +
  xlab("\nMonth of year (2021 - 2022)") +
  ylab("Depth (m below sea level)")

shark_daily_depth_plot
```

# Separately

# Max only
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_maxdepth_plot <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  theme_bjc() +
  scale_y_reverse(limits = c(201, 0)) +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  xlab("Date") +
  ylab("Max daily depth (m)")

# View
shark_daily_maxdepth_plot
# Save
#ggsave(filename = "output/shark_daily_maxdepth_plot.pdf", plot = shark_daily_maxdepth_plot, 
#       device = cairo_pdf, width = 140, height = 70, units = "mm", dpi = 300)
```


# Max and min 
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_maxmindepth_plot <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +  
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  theme_bjc() +
  scale_y_reverse(limits = c(201, 0)) +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  xlab("Date") +
  ylab("Daily depth (m)")

# View
shark_daily_maxmindepth_plot

# Save
#ggsave(filename = "output/shark_daily_maxmindepth_plot.pdf", plot = shark_daily_maxmindepth_plot, 
 #      device = cairo_pdf, width = 140, height = 70, units = "mm", dpi = 300)
```

# Mean
```{r fig.width = 140/25.4, fig.height = 70/25.4}
shark_daily_meandepth_plot <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  theme_bjc() +
  scale_y_reverse(limits = c(51, 0)) +
  scale_x_date(date_breaks = "1 months", date_labels = '%b') +
  xlab("Date") +
  ylab("Daily depth (m)")

# View
shark_daily_meandepth_plot

# Save
#ggsave(filename = "output/shark_daily_meandepth_plot.pdf", plot = shark_daily_meandepth_plot, 
 #      device = cairo_pdf, width = 140, height = 70, units = "mm", dpi = 300)
```




## Plot max depths vs SST
```{r fig.width = 140/25.4, fig.height = 70/25.4}
max_dep_SST_fig <- cowplot::plot_grid(sst_2021_2022, shark_daily_maxdepth_plot, nrow = 2, rel_heights = c(0.6, 1))

# Save
#ggsave(filename = "output/MaxDep_SST_plot.pdf", plot = MaxDep_SST_plot, 
#       device = cairo_pdf, width = 140, height = 40, units = "mm", dpi = 300)
```

## Can use a quick regression to see the relationship
```{r fig.width = 70/25.4, fig.height = 70/25.4}
shark_SST_maxdep_lm_plot <- shark_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_max_depth)) +
  #geom_smooth(aes(x = daily_max_temp, y = daily_max_depth), method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_smooth(aes(x = SST, y = daily_max_depth), method = "glm") +
  theme_bjc() +
  ylab("Daily maximum depth (mbsl)") + 
  xlab("SST")
shark_SST_maxdep_lm_plot
```

It looks like the 5°C difference (summer vs winter) leads to a 50m difference in maximum daily depths, from 60 - 120m

```{r}
#ggsave(filename = "output/shark_SST_maxdep_lm_plot.pdf", plot = shark_SST_maxdep_lm_plot,  width = 100, height = 100, units = "mm", dpi = 300)
```

# Combine figs
```{r fig.height=60/25.4, fig.width=180/25.4}
daily_max_dep_sum_fig <- 
  ggarrange(max_dep_SST_fig, shark_SST_maxdep_lm_plot, ncol = 2, widths = c(1.6,1))
daily_max_dep_sum_fig
```

# And save
```{r}
#ggsave(filename = "output/daily_max_dep_sum_fig.pdf", plot = daily_max_dep_sum_fig, width = 180, height = 60, units = "mm", dpi = 300)
```



## Final plot - SST / Shark temps / Max depth

```{r fig.width = 140/25.4, fig.height = 100/25.4}
temp_plot_plain <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_line(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_bjc() +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +    xlab("\nDate") + 
  ylab("Shark daily temps\n(°C)")

depth_plot_plain <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  theme_bjc() +
  scale_y_reverse() +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme(axis.text.x = element_text(angle = 45)) +
  xlab("Date") +
  ylab("Shark daily depths\n(m below sealevel)")

combined_plot <- 
    ggarrange(sst_2021_2022, temp_plot_plain, depth_plot_plain, nrow = 3, heights = c(0.5, 1, 1.5))

# View
combined_plot

#ggsave(filename = "output/combined_plot.pdf", plot = combined_plot, width = 161, height = 280, units = "mm", dpi = 300)
```


```{r}
# END #
```

