---
title: "Individual shark daily depth and temperature data"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document: 
   code_folding: show
   collapse: no
   fig_caption: yes
   fig_height: 10
   fig_width: 10
   highlight: textmate
   theme: spacelab
   toc: yes
   toc_float: yes
editor_options: 
  chunk_output_type: inline
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
---


```{r}
#knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
rm(list=ls()) # Clear out environment if necessary
source("code/packages.R")
```

# Load shark T/P data
```{r}
load(file = "data/Rdata/shark_tp_data.Rdata")
```

# Load SST data
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
sourceRmd("code/00_wrangle/wrangling_SSTdata.Rmd")
```

# Individual daily temp summary
Going to add in SST data at this point also
```{r}
shark_individual_daily_temps <- 
  shark_tp_data %>% 
  #mutate(date = date(detection_timestamp)) %>% 
  filter(!is.na(Temp)) %>% 
  group_by(date, transmitter_id, Sex, TL) %>% 
  summarise(daily_max_temp = max(Temp),
            daily_min_temp = min(Temp),
            daily_mean_temp = mean(Temp),
            daily_temp_sd = sd(Temp)) %>% 
  full_join(SST %>% filter(date > date("2021-10-27"))) # Filter out dates earlier than the first detection 
#shark_individual_daily_temps
```


## Individual daily depth summary
Going to add in SST data, Sex and TL at this point also as we'll need for modelling shortly
```{r}
shark_individual_daily_depths <- 
  shark_tp_data %>% 
  #mutate(date = date(detection_timestamp)) %>% 
  filter(!is.na(Depth)) %>% 
  group_by(date, transmitter_id, Sex, TL) %>% 
  summarise(daily_max_depth = max(Depth),
            daily_min_depth = min(Depth),
            daily_mean_depth = mean(Depth),
            daily_depth_sd = sd(Depth),
            daily_movement = daily_max_depth - daily_min_depth) %>% 
  full_join(SST %>% filter(date > date("2021-10-27"))) # Filter out dates earlier than the first detection 
#shark_individual_daily_depths
```

## Combine the individual temp and depth dfs into one and add in season (for modelling)
```{r}
shark_individual_daily_td <- 
 shark_individual_daily_depths %>% 
  full_join(shark_individual_daily_temps) %>% 
  mutate(Month = as_factor(month(date, label = TRUE, abbr = TRUE)),
         Season = factor(case_when(Month %in% c("Dec" , "Jan", "Feb") ~ "Summer",
                                   Month  %in% c("Mar" , "Apr", "May") ~ "Autumn",
                                   Month  %in% c("Jun", "Jul", "Aug") ~ "Winter",
                                   Month  %in% c("Sep" , "Oct", "Nov") ~ "Spring"))) %>% 
  ungroup()
```

## Note that season automatically gets levelled by N hemisphere!!??
```{r}
shark_individual_daily_td  %$% 
  summary(Season)
```

# Save 
```{r}
save(shark_individual_daily_td, file = "data/Rdata/shark_individual_daily_td.Rdata")
```


## Plots
# Temps
```{r}
shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_max_temp), colour = "red") +
  geom_smooth(aes(x = SST, y = daily_max_temp), colour = "black", method = "lm") 

shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_mean_temp), colour = "grey40") +
  geom_smooth(aes(x = SST, y = daily_mean_temp), colour = "black", method = "lm") 

shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_min_temp), colour = "blue") +
  geom_smooth(aes(x = SST, y = daily_min_temp), colour = "black", method = "lm") 

shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_temp_sd), colour = "grey40") +
  geom_smooth(aes(x = SST, y = daily_temp_sd), colour = "black", method = "glm") 

shark_individual_daily_td %>% 
ggplot() +
  geom_histogram(aes(x = daily_temp_sd), colour = "grey40") 
```

# Depth
```{r}
shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_max_depth), colour = "grey60") +
  geom_smooth(aes(x = SST, y = daily_max_depth), colour = "black", method = "lm") +
  scale_y_reverse()

shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_mean_depth), colour = "grey20") +
  geom_smooth(aes(x = SST, y = daily_mean_depth), colour = "black", method = "lm") +
  scale_y_reverse()

shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_min_depth), colour = "grey10") +
  geom_smooth(aes(x = SST, y = daily_min_depth), colour = "black", method = "lm") +
  scale_y_reverse()

shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_movement), colour = "grey60") +
  geom_smooth(aes(x = SST, y = daily_movement), colour = "black", method = "lm") 

shark_individual_daily_td %>% 
ggplot() +
  geom_point(aes(x = SST, y = daily_depth_sd), colour = "grey40") +
  geom_smooth(aes(x = SST, y = daily_depth_sd), colour = "black", method = "glm") +
  scale_y_reverse()

shark_individual_daily_td %>% 
ggplot() +
  #geom_histogram(aes(x = daily_depth_sd), colour = "grey40") 
  #geom_histogram(aes(x = daily_max_depth), colour = "grey40") 
  #geom_histogram(aes(x = daily_mean_depth), colour = "grey40") 
  geom_histogram(aes(x = daily_min_depth), colour = "grey40") 


```



```{r}
shark_individual_daily_td %>% 
  filter(!is.na(daily_max_depth),
         !is.na(daily_movement)) %>% 
  group_by(transmitter_id) %>% 
  summarise(mean_daily_max_depth = mean(daily_max_depth),
            mean_daily_movt = mean(daily_movement))

```




```{r fig.width=16, fig.height=6}
shark_individual_daily_temps_plot <- 
shark_individual_daily_td %>% 
ggplot() +
  geom_line(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_smooth(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_line(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_smooth(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  geom_smooth(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #geom_line(aes(x = week_date, y = weekly_rolling_mean), data = SSTcomb, colour = "black") +
  #geom_smooth(aes(x = week_date, y = weekly_rolling_mean), data = SSTcomb, colour = "black") +
  geom_line(aes(x = date, y = SST), colour = "black") +
  #geom_line(aes(x = Month, y = SST), data = AMM_for_plot, colour = "black") +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  #annotate("text", x = as.Date("2021-11-01"), y = 30.5, label = "a") +
  xlab("\nDate") + 
  ylab("Temp (°C)") +
  facet_wrap(facets = "transmitter_id", ncol = 3)
shark_individual_daily_temps_plot
```



## Plot daily depths
```{r fig.width=16, fig.height=6}
shark_individual_daily_depth_plot <- 
shark_individual_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  #geom_smooth(aes(x = date, y = daily_min_depth), colour = "blue",  method = "gam", formula = y ~ s(x, bs = 'cs')) + # method = "gam" implements mgcv
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_smooth(aes(x = date, y = daily_mean_depth), colour = "dodgerblue",  method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "red") + # Basic smooth, rather than gam
  theme_minimal() +
  scale_y_reverse() +
  #scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  #theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +
  xlab("\nDate") + 
  ylab("Depth \n (m below sealevel)") +
  facet_wrap(facets = "transmitter_id", ncol = 3)

shark_individual_daily_depth_plot
```


# Explore individual depth responses to higher SSTs
```{r fig.width = 10, fig.height = 16}
#shark_td_lm_individ_plot <- 
shark_individual_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_max_depth)) +
  #geom_smooth(aes(x = daily_max_temp, y = daily_max_depth), method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_smooth(aes(x = SST, y = daily_max_depth), method = "lm") +
  theme_minimal() +
  ylab("Daily maximum depth (mbsl)") + 
  xlab("SST") +
  facet_wrap(facets = "transmitter_id", ncol = 3)
#shark_td_lm_individ_plot
```



```{r}
#END
```


