---
title: "Shark data: daily"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_width: 161/25.4
    fig_height: 100/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("code/packages.R")
rm(list=ls()) # Clear out environment if necessary
```

# Load data 
## Shark T/P data
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
load(file = "data/Rdata/shark_tp_data.Rdata")
# Or, if you also want all the associated files loaded:
#source(knitr::purl("code/00_wrangle/wrangling_sensordata.Rmd", quiet = TRUE))
```

## SST
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
load(file = "data/RData/SST.Rdata")
# Or, if you also want all the associated files loaded:
#source(knitr::purl("code/00_wrangle/wrangling_SSTdata.Rmd", quiet = TRUE))
```


# Shark depth - daily summary
```{r}
shark_daily_depths <- 
  shark_tp_data %>% 
  filter(!is.na(Depth)) %>% 
  group_by(date) %>% 
  summarise(daily_max_depth = max(Depth),
            daily_min_depth = min(Depth),
            daily_mean_depth = mean(Depth),
            daily_depth_sd = sd(Depth)) %>% 
  mutate(daily_min_depth = if_else(daily_min_depth <0, 0, daily_min_depth))
#shark_daily_depths
#save(shark_daily_depths, file = "data/Rdata/shark_daily_depths.Rdata")
```

# Shark temps - daily summary
```{r}
shark_daily_temps <- 
  shark_tp_data %>% 
  filter(!is.na(Temp)) %>% 
  group_by(date) %>% 
  summarise(daily_max_temp = max(Temp),
            daily_min_temp = min(Temp),
            daily_mean_temp = mean(Temp),
            daily_temp_sd = sd(Temp))
#shark_daily_temps
#save(shark_daily_temps, file = "data/Rdata/shark_daily_temps.Rdata")
```


# Combine
```{r}
shark_daily_td <- 
  SST %>% 
  full_join(shark_daily_depths) %>% 
  full_join(shark_daily_temps) %>% 
  filter(date > as.Date("2021-10-29"))
```

## Daily temps (all) vs SST
```{r}
shark_daily_temps_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 1) +
  geom_line(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_smooth(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_line(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_smooth(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  geom_smooth(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  #annotate("text", x = as.Date("2021-11-01"), y = 30.5, label = "a") +
  xlab("\nDate") + 
  ylab("Temp (°C)")
# View
shark_daily_temps_plot
# Save
ggsave(filename = "output/shark_daily_temps_plot.pdf", plot = shark_daily_temps_plot, width = 161, height = 100, units = "mm", dpi = 300)
```

# Daily maximumm temperatures track closely with SST
```{r}
shark_daily_maxtemp_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.5) +
  geom_line(aes(x = date, y = daily_max_temp),  colour = "red") +
  #geom_smooth(aes(x = date, y = daily_max_temp), colour = "red") +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("\nDate") + 
  ylab("Temp (°C)")

# View
shark_daily_maxtemp_plot
# Save
ggsave(filename = "output/shark_daily_maxtemp_plot.pdf", plot = shark_daily_maxtemp_plot, width = 161, height = 100, units = "mm", dpi = 300)
```


# However, during summer months sharks were able to negatively thermoregulate mean body temperatures despite high SSTs:
```{r}
shark_daily_meantemp_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.5) +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #geom_smooth(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #geom_vline(xintercept = as.Date("2022-01-19"), colour = "red", lwd = 1) +
  geom_segment(aes(x = as.Date("2022-01-19"), xend = as.Date("2022-01-19"), y = 29.6, yend = 28.4), colour = "blue",
                  arrow = arrow(length = unit(0.2, "cm"))) +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("\nDate") + 
  ylab("Temp (°C)")

# View
shark_daily_meantemp_plot
# Save
ggsave(filename = "output/shark_daily_meantemp_plot.pdf", plot = shark_daily_meantemp_plot, width = 161, height = 100, units = "mm", dpi = 300)
```

# The maximum daily downward thermo-regulation was as much as 2°C (on 19 Jan 2022), when SST was around 30°C. Max depth recorded on this date was 122m. (Potentially could investigate what this would come out as at individual level?)
```{r}
shark_daily_temp_diff <- 
  shark_daily_td %>% 
  mutate(temp_diff = daily_max_temp - daily_mean_temp)

shark_daily_temp_diff %>% 
  #summarise(max_temp_diff = max(temp_diff))
  filter(temp_diff == max(temp_diff))
```


## When plotted over the course of the year, it is obvious that these diffs are higher in the summer
```{r}
shark_daily_temp_diff %>% 
  ggplot() +
  geom_line(aes(x = date, y = temp_diff)) +
  geom_smooth(aes(x = date, y = temp_diff)) +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("\nDate") + 
  ylab("Daily temp diff\n(max vs mean daily temp, °C)")
```



## Relationship between SST


## Minimum daily temperatures were consistent throughout the year and did not depend on SST (25.3°C ± 0.6°C SD)
### Calculate avg and SD
```{r}
shark_daily_td %>% 
  summarise(avg_daily_min_temp = mean(daily_min_temp),
            sd_daily_min_temp = sd(daily_min_temp))
```

### Plot relationship between SST and daily min temps (there isn't one)
```{r}
shark_SST_mintemp_lm_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_min_temp)) +
  geom_smooth(aes(x = SST, y = daily_min_temp), method = "glm") +
  theme_minimal() +
  ylab("Daily min temp (°C)") + 
  xlab("SST (°C)")

# View  
shark_SST_mintemp_lm_plot

# Save
ggsave(filename = "output/shark_SST_mintemp_lm_plot.pdf", plot = shark_SST_mintemp_lm_plot, width = 100, height = 100, units = "mm", dpi = 300)
```


```{r}
shark_daily_mintemp_plot <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.5) +
  geom_line(aes(x = date, y = daily_min_temp),  colour = "blue") +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("\nDate") + 
  ylab("Temp (°C)")

# View
shark_daily_mintemp_plot
# Save
ggsave(filename = "output/shark_daily_mintemp_plot.pdf", plot = shark_daily_mintemp_plot, width = 161, height = 100, units = "mm", dpi = 300)

```

# Combine figs
```{r}
daily_min_temp_sum_fig <- 
  ggarrange(shark_daily_mintemp_plot, shark_SST_mintemp_lm_plot, ncol = 2, widths = c(1.6,1))
```

# And save
```{r}
ggsave(filename = "output/daily_min_temp_sum_fig.pdf", plot = daily_min_temp_sum_fig, width = 180, height = 60, units = "mm", dpi = 300)
```





## Downward thermoregulation appears to correspond with maximum daily depths attained - depth movements used in order to thermoregulate
```{r}
shark_daily_temp_diff %>% 
  ggplot() +
  geom_point(aes(x = temp_diff, y = daily_max_depth)) +
  #geom_smooth(aes(x = daily_max_temp, y = daily_max_depth), method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_smooth(aes(x = temp_diff, y = daily_max_depth), method = "lm") +
  theme_minimal() +
  ylab("Daily maximum depth (mbsl)") + 
  xlab("Daily temperature difference (°C)")
```

... so let's explore depths a bit more


## Plot daily depths
```{r}
shark_daily_depth_plot <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  #geom_smooth(aes(x = date, y = daily_min_depth), colour = "blue",  method = "gam", formula = y ~ s(x, bs = 'cs')) + # method = "gam" implements mgcv
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_smooth(aes(x = date, y = daily_mean_depth), colour = "dodgerblue",  method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "red") + # Basic smooth, rather than gam
  theme_minimal() +
  scale_y_reverse() +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme(axis.text.x = element_text(angle = 45)) +
  #theme(axis.text.x = element_blank()) +
  #theme(axis.title.x = element_blank()) +
  #theme(axis.ticks.x = element_blank()) +
  xlab("Date") +
  ylab("Depth \n (m below sealevel)")

shark_daily_depth_plot
```


```{r}
shark_daily_maxdepth_plot <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  theme_minimal() +
  scale_y_reverse() +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme(axis.text.x = element_text(angle = 45)) +
  #theme(axis.text.x = element_blank()) +
  #theme(axis.title.x = element_blank()) +
  #theme(axis.ticks.x = element_blank()) +
  xlab("Date") +
  ylab("Max daily depth \n (m below sealevel)")

# View
shark_daily_maxdepth_plot
# Save
ggsave(filename = "output/shark_daily_maxdepth_plot.pdf", plot = shark_daily_maxdepth_plot, width = 161, height = 100, units = "mm", dpi = 300)
```


## Max depths vs SST
```{r}
SST_plot_plain <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = SST), colour = "grey15", lwd = 0.5) +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +  
  xlab("\nDate") + 
  ylab("SST\n(°C)")

(MaxDep_SST_plot <- SST_plot_plain/shark_daily_maxdepth_plot)

# Save
ggsave(filename = "output/MaxDep_SST_plot.pdf", plot = MaxDep_SST_plot, width = 161, height = 100, units = "mm", dpi = 300)
```

## Can use a quick regression to see the relationship
```{r}
shark_SST_maxdep_lm_plot <- 
shark_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_max_depth)) +
  #geom_smooth(aes(x = daily_max_temp, y = daily_max_depth), method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_smooth(aes(x = SST, y = daily_max_depth), method = "glm") +
  theme_minimal() +
  ylab("Daily maximum depth (mbsl)") + 
  xlab("SST")
shark_SST_maxdep_lm_plot
```

It looks like the 5°C difference (summer vs winter) leads to a 50m difference in maximum daily depths, from 60 - 120m

```{r}
ggsave(filename = "output/shark_SST_maxdep_lm_plot.pdf", plot = shark_SST_maxdep_lm_plot,  width = 100, height = 100, units = "mm", dpi = 300)
```

# Combine figs
```{r fig.height=60/25.4, fig.width=180/25.4}
daily_max_dep_sum_fig <- 
  ggarrange(MaxDep_SST_plot, shark_SST_maxdep_lm_plot, ncol = 2, widths = c(1.6,1))
daily_max_dep_sum_fig
```

# And save
```{r}
ggsave(filename = "output/daily_max_dep_sum_fig.pdf", plot = daily_max_dep_sum_fig, width = 180, height = 60, units = "mm", dpi = 300)
```



## Final plot - SST / Shark temps / Max depth

```{r}
temp_plot_plain <- 
  shark_daily_td %>% 
  ggplot() +
  geom_line(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_line(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +    xlab("\nDate") + 
  ylab("Shark daily temps\n(°C)")

depth_plot_plain <- 
shark_daily_td %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  theme_minimal() +
  scale_y_reverse() +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme(axis.text.x = element_text(angle = 45)) +
  xlab("Date") +
  ylab("Shark daily depths\n(m below sealevel)")

combined_plot <- 
    ggarrange(SST_plot_plain, temp_plot_plain, depth_plot_plain, nrow = 3, heights = c(0.5, 1, 1))

# View
combined_plot

ggsave(filename = "output/combined_plot.pdf", plot = combined_plot, width = 161, height = 280, units = "mm", dpi = 300)
```




```{r}
# END #
```

