---
title: "Individual shark daily depth and temperature patterns"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
output_dir: "docs"
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
monofont: DejaVu Sans Mono
classoption: a4paper
---


```{r}
knitr::opts_chunk$set(echo = TRUE)
source("code/packages.R")
#rm(list=ls()) # Clear out environment if necessary
```

# Load shark T/P data
```{r}
load(file = "data/Rdata/shark_tp_data.Rdata")
```

# Load SST data
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
source(knitr::purl("code/00_wrangle/wrangling_SSTdata.Rmd", quiet = TRUE))
```


# Shark temps - individual daily summary
Going to add in SST data at this point also
```{r}
shark_individual_daily_temps <- 
  shark_tp_data %>% 
  mutate(date = date(detection_timestamp)) %>% 
  filter(!is.na(Temp)) %>% 
  group_by(date, transmitter_id, Sex, TL) %>% 
  summarise(daily_max_temp = max(Temp),
            daily_min_temp = min(Temp),
            daily_mean_temp = mean(Temp),
            daily_temp_sd = sd(Temp)) %>% 
  full_join(SST)
shark_individual_daily_temps
```

## Plot
```{r fig.width=16, fig.height=6}
shark_individual_daily_temps_plot <- 
shark_individual_daily_temps %>% 
ggplot() +
  geom_line(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_smooth(aes(x = date, y = daily_max_temp), colour = "red") +
  geom_line(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_smooth(aes(x = date, y = daily_min_temp), colour = "blue") +
  geom_line(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  geom_smooth(aes(x = date, y = daily_mean_temp), colour = "grey40") +
  #geom_line(aes(x = week_date, y = weekly_rolling_mean), data = SSTcomb, colour = "black") +
  #geom_smooth(aes(x = week_date, y = weekly_rolling_mean), data = SSTcomb, colour = "black") +
  geom_line(aes(x = date, y = SST), data = SST, colour = "black") +
  #geom_line(aes(x = Month, y = SST), data = AMM_for_plot, colour = "black") +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  #annotate("text", x = as.Date("2021-11-01"), y = 30.5, label = "a") +
  xlab("\nMonth of year") + 
  ylab("Temp (°C)") +
  facet_wrap(facets = "transmitter_id")
shark_individual_daily_temps_plot
```


# Shark temps - individual weekly summary
-> Might not need this, given we have daily SST values
```{r}
shark_individual_weekly_temps <- 
  shark_tp_data %>% 
  mutate(week_date = date(floor_date(detection_timestamp, "week"))) %>%
  filter(!is.na(Temp)) %>% 
  group_by(week_date, transmitter_id, Sex, TL) %>% 
  summarise(weekly_max_temp = max(Temp),
            weekly_min_temp = min(Temp),
            weekly_mean_temp = mean(Temp))
shark_individual_weekly_temps
```

# Plot individual weekly temps (not really different from daily depths but potentially easier to visualise/interpret)
```{r fig.width=16, fig.height=6}
shark_individual_weekly_temps_plot <- 
shark_individual_weekly_temps %>% 
  ggplot() +
  geom_line(aes(x = week_date, y = weekly_min_temp), colour = "blue") +
  #geom_smooth(aes(x = week_date, y = weekly_min_temp), colour = "blue", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  geom_smooth(aes(x = week_date, y = weekly_min_temp), colour = "blue", method = "loess", level = 0.95) +
  geom_line(aes(x = week_date, y = weekly_mean_temp), colour = "grey30") +
  geom_smooth(aes(x = week_date, y = weekly_mean_temp), colour = "grey30", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  geom_line(aes(x = week_date, y = weekly_max_temp), colour = "red") +
  geom_smooth(aes(x = week_date, y = weekly_max_temp), colour = "red", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  geom_line(aes(x = date, y = SST), data = SST, colour = "black") +
  scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  xlab("Date") +
  ylab("Averaged weekly depths \n (m below sealevel)") +
  facet_wrap(facets = "transmitter_id")

shark_individual_weekly_temps_plot
```


# Shark depths

## Individual daily depth summary
Going to add in SST data, Sex and TL at this point also as we'll need for modelling shortly
```{r}
shark_individual_daily_depths <- 
  shark_tp_data %>% 
  mutate(date = date(detection_timestamp)) %>% 
  filter(!is.na(Depth)) %>% 
  group_by(date, transmitter_id, Sex, TL) %>% 
  summarise(daily_max_depth = max(Depth),
            daily_min_depth = min(Depth),
            daily_mean_depth = mean(Depth),
            daily_depth_sd = sd(Depth))%>% 
  mutate(daily_min_depth = if_else(daily_min_depth <0, 0, daily_min_depth),
         daily_mean_depth = if_else(daily_mean_depth <0, 0, daily_mean_depth),
         daily_max_depth = if_else(daily_max_depth <0, 0, daily_max_depth)) %>%    # Depths cannot be out of the water, so must be instrument/calibration glitch
  full_join(SST)
shark_individual_daily_depths
```

## Plot daily depths
```{r fig.width=16, fig.height=6}
shark_individual_daily_depth_plot <- 
shark_individual_daily_depths %>%   
  ggplot() +
  geom_line(aes(x = date, y = daily_min_depth), colour = "skyblue") +
  #geom_smooth(aes(x = date, y = daily_min_depth), colour = "blue",  method = "gam", formula = y ~ s(x, bs = 'cs')) + # method = "gam" implements mgcv
  geom_line(aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_smooth(aes(x = date, y = daily_mean_depth), colour = "dodgerblue",  method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_line(aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_smooth(aes(x = date, y = daily_max_depth), colour = "midnightblue",  method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  #geom_smooth(aes(x = date, y = daily_max_depth), colour = "red") + # Basic smooth, rather than gam
  theme_minimal() +
  scale_y_reverse() +
  #scale_x_date(date_breaks = "1 months", date_labels = '%Y-%m') +
  #theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) +
  #xlab("Date") +
  ylab("Depth \n (m below sealevel)") +
  facet_wrap(facets = "transmitter_id", ncol = 3)

shark_individual_daily_depth_plot
```


# Weekly summary - by individual
```{r}
shark_individual_weekly_depths <- 
shark_tp_data %>% 
  mutate(week_date = date(floor_date(detection_timestamp, "week"))) %>% 
  filter(!is.na(Depth)) %>% 
  group_by(week_date, transmitter_id, Sex, TL) %>% 
  summarise(weekly_max_depth = max(Depth),
            weekly_min_depth = min(Depth),
            weekly_mean_depth = mean(Depth),
            weekly_depth_sd = sd(Depth))%>% 
  mutate(weekly_min_depth = if_else(weekly_min_depth <0, 0, weekly_min_depth))

shark_individual_weekly_depths
```

## Plot individual weekly depths
```{r fig.width=16, fig.height=6}
shark_individual_weekly_depth_plot <- 
shark_individual_weekly_depths %>% 
  ggplot() +
  geom_line(aes(x = week_date, y = weekly_min_depth), colour = "blue") +
  geom_line(aes(x = week_date, y = weekly_mean_depth), colour = "grey30") +
  geom_smooth(aes(x = week_date, y = weekly_mean_depth), colour = "grey30", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  geom_line(aes(x = week_date, y = weekly_max_depth), colour = "red") +
  geom_smooth(aes(x = week_date, y = weekly_max_depth), colour = "red", method = "gam", formula = y ~ s(x, bs = 'cs'), level = 0.95) +
  scale_y_reverse() +
  theme_minimal() +
  xlab("Date") +
  ylab("Averaged weekly depths \n (m below sealevel)") +
  facet_wrap(facets = "transmitter_id")

shark_individual_weekly_depth_plot
```



## Combine the individual temp and depth dfs into one and add in season (for modelling)
```{r}
shark_individual_daily_td <- 
 shark_individual_daily_depths %>% 
  full_join(shark_individual_daily_temps) %>% 
  mutate(Season = factor(case_when(
    date >= date("2022-01-01") & date<= date("2022-03-31") ~ "Summer",
    date >= date("2022-07-01") & date<= date("2022-09-30") ~ "Winter",
    date < date("2022-07-01") & date > date("2022-03-31") ~ "Autumn",
    date < date("2022-12-31")  ~ "Spring",
    date > date("2022-09-30") ~ "Spring"))) %>% 
  filter(date > date("2021-10-27"))
```

```{r}
shark_td_lm_individ_plot <- 
shark_individual_daily_td %>% 
  ggplot() +
  geom_point(aes(x = SST, y = daily_max_depth)) +
  #geom_smooth(aes(x = daily_max_temp, y = daily_max_depth), method = "gam", formula = y ~ s(x, bs = 'cs')) +
  geom_smooth(aes(x = SST, y = daily_max_depth), method = "lm") +
  theme_minimal() +
  ylab("Daily maximum depth (mbsl)") + 
  xlab("SST") +
  facet_wrap(facets = "transmitter_id")
shark_td_lm_individ_plot
```

# Save 
```{r}
save(shark_individual_daily_td, file = "data/Rdata/shark_individual_daily_td.Rdata")
```


```{r}
#END
```


