---
title: "Shark daily depths ~ SST models"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


```{r include=FALSE}
source("code/packages.R")
source("code/extras.R")
```

# Load data
```{r}
# Using the shark grouped Rmd
sourceRmd("code/01_eda/shark_daily_grouped.Rmd")
shark_daily_td
```


# We need models for:

5. Max daily depth ~ SST:               Gaussian GLM
6. Mean daily depth ~ SST
7. Min daily depth ~ SST
8. Variability in daily depth ~ SST

## Check distributions to get an idea of best error dist for models
```{r}
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_max_depth)) # Prob gaussian
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_mean_depth)) # Almost bi-model but will try gaussian
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_min_depth)) # Looks zero-inflated but will check later on in DHARMa - prob NB
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_depth_sd)) #
```

All look approximately normal, apart from min depth which is almost always 0 (i.e. at the surface)


# First, need a grid of SST values to use for plotting all models at the end
```{r}
SSTgrid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))
```



##### Model 5: Daily Max Depth ~ SST #####

## Model formula
```{r}
mod5 <- glm(daily_max_depth ~ SST, data = shark_daily_td, family = gaussian)
mod5b <- gam(daily_max_depth ~ s(SST, bs = "cs", k = 3), data = shark_daily_td, family = gaussian, method = "ML")
```

## Model validation
```{r}
# Dharma resids
mod5 %>% simulateResiduals(plot = TRUE) # Look ok 
mod5 %>% performance::check_outliers()
mod5 %>% performance::r2()

AIC(mod5, mod5b) # GAM adds nothing
```

## Results
```{r}
# broom::tidy
(mod5results <- mod5 %>% tidy(conf.int=TRUE))

# Manually - per 1°C increase SST
mod5 %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30.5, 25.5)))) %>% 
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()

```

# So for every 1° increase in SST the sharks max daily depth on average increases by 14m 95%CI [12.4-15.5]

# 25.5 - 30.5 C SST contrast



## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod5_plot <-
  mod5 %>% 
  emmeans(~SST, at = SSTgrid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  ggplot(aes(y = emmean, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_max_depth, x = SST), fill = "midnightblue", col = "black", alpha = 0.5, pch = 21) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), fill = "midnightblue", alpha=0.6) +
  geom_line(color = "black") +
  ylab("Daily max depth (m)") + 
  xlab("SST (°C)") +
  scale_x_continuous(limits = c(25.6, 30.6), n.breaks = 6) +
  theme_bjc() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) 

mod5_plot
```


```{r fig.height = 56/25.4, fig.width = 56/25.4}
ggsave(mod5_plot, file = "output/mod5_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```


##### Model 6: Daily Mean Depth ~ SST #####

## Model formula - Gamma GAM best fit
```{r}
mod6a <- glm(daily_mean_depth ~ SST, data = shark_daily_td)
mod6b <- glm(daily_mean_depth ~ SST, data = shark_daily_td, family = Gamma)
mod6c <- gam(daily_mean_depth ~ s(SST, bs = "cs", k = 3), data = shark_daily_td, family = Gamma, method = "ML")
```

## Model validation
```{r}
# Dharma resids
mod6a %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod6a %>% performance::check_outliers()
mod6a %>% performance::r2()

mod6b %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod6b %>% performance::check_outliers()
mod6b %>% performance::r2()

mod6c %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod6c %>% performance::check_outliers()
mod6c %>% performance::r2()

AIC(mod6a, mod6b, mod6c)
```

```{r}
mod6 <- mod6c
```


## Results
```{r}
# broom::tidy
(mod6results <- mod6 %>% tidy(conf.int=TRUE, exponentiate = TRUE))
mod6 |> summary()

# Manually
mod6 %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30.5, 30, 29, 28, 27, 26, 25.5))), type = "response") %>% 
  regrid() %>% 
 # pairs() %>% 
  confint() %>% 
  as_tibble()

# Contrast across SST range
mod6 %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30.5, 25.5))), type = "response") %>% 
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()

# Between 25.5 and 26.5
mod6 %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(26.5, 25.5))), type = "response") %>% 
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()

# Between 29.5 and 30.5
mod6 %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30.5, 29.5))), type = "response") %>% 
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()

```
Mean daily depth at lowest SST (25.5°C) is 15m and at highest SST (30.5°C) it is over double - 41m.
Mean daily depth at 26°C SST is 15m and at 30°C it is 33m. And not a linear trend:
1°C increase in SST at low range (25.5-26.5) results in 1.5m greater mean daily depth while
1°C increase in SST at high range (29.5-30.5) results in a 13m greater mean daily depth - nearly 10x the response.


## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod6_plot <- 
  mod6 %>% 
  emmeans(~SST, at = SSTgrid, type = "response") %>% 
  #regrid(transform = TRUE, ) %>% 
  as_tibble() %>% 
  ggplot(aes(y = response, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_mean_depth, x = SST), fill = "dodgerblue", col = "black", alpha = 0.6, pch = 21) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), fill = "dodgerblue", alpha=0.6) +
  geom_line(colour = "black") +
  ylab("Daily mean depth (m)") + 
  xlab("SST (°C)") +
  scale_x_continuous(limits = c(25.6, 30.6), n.breaks = 6) +
  theme_bjc() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) 

mod6_plot
```


# Save
```{r fig.height = 56/25.4, fig.width = 56/25.4}
ggsave(mod6_plot, file = "output/mod6_plot.pdf", device = cairo_pdf, width = 56, height = 56, units = "mm", dpi = 300)
```



## Add to time-series data if required
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
#(daily_mean_depth_model_plot <- ggarrange(shark, mod6_plot, ncol = 2, widths = c(2,1)))
#ggsave(daily_mean_depth_model_plot, file = "output/daily_mean_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)
```



##### Model 7: Daily Min Depth ~ SST #####

## Model formula - have to use negbinom due to all the 0s
```{r}
mod7a <- glm(daily_min_depth ~ SST, data = shark_daily_td) 
mod7b <- glm(daily_min_depth ~ SST, data = shark_daily_td, family = negative.binomial(2)) # nb capacity comes from MASS package, so have to have this loaded
```

## Model validation
```{r}
# Dharma resids
mod7a %>% simulateResiduals(plot = TRUE) # This is why you have to use nb!
mod7b %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod7b %>% performance::check_outliers()
mod7b %>% testZeroInflation
mod7b %>% performance::r2()
```

```{r}
mod7 <- mod7b
```


## Results
```{r}
(mod7results <- mod7 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod7_plot <- 
  mod7 %>% 
  emmeans(~SST, at = SSTgrid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  ggplot(aes(y = response, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_min_depth, x = SST), fill = "skyblue", col = "black", alpha = 0.6, pch = 21) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), fill = "skyblue", alpha=0.3) +        # Had to change the ymin and max names as the model conducted in MASS
  geom_line(colour = "black") +
  ylab("Daily min depth (m)") + 
  xlab("SST (°C)") +
  scale_x_continuous(limits = c(25.6, 30.6), n.breaks = 6) +
  theme_bjc() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) 


mod7_plot
```


```{r fig.height = 56/25.4, fig.width = 56/25.4}
ggsave(mod7_plot, file = "output/mod7_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```



## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
#(daily_min_depth_model_plot <- ggarrange(MinDep_SST_plot, mod7_plot, ncol = 2, widths = c(2,1)))
#ggsave(daily_min_depth_model_plot, file = "output/daily_min_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)
```









##### Model 8: Variability in depth use ~ SST #####

# Depth variability Pt I - SD

## SD first
Need a GAM for this
```{r}
mod8a <- gam(daily_depth_sd ~ s(SST, bs = "cs", k = 3), data = shark_daily_td, method = "ML")
mod8b <- gam(daily_depth_sd ~ s(SST, bs = "cs", k = 3), data = shark_daily_td, family = tw, method = "ML")
```

## Validation and Information criteria
```{r}
# Dharma resids
mod8a %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod8b %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
# AIC
AIC(mod8a, mod8b)
```
Of the gams, the tw is the better fit. 


```{r}
mod8_sd <- mod8b
```


# Plot
```{r fig.height = 100/25.4, fig.width = 100/25.4}
mod8_sd_plot <- 
  mod8_sd %>% 
  emmeans(~SST, at = SSTgrid) %>% 
  regrid() %>% 
  as_tibble() %>% 
    ggplot(aes(y = response, x = SST)) +
    geom_point(data=shark_daily_td,  aes(y=daily_depth_sd, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.7) +
    geom_line() +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Daily mean depth S.D (m)") + 
    xlab("SST (°C)") +
    theme_bjc() +
    theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9))
mod8_sd_plot
```

# Results

```{r}
mod8_sd %>%  summary()
```




# Planned contrast SST extremes
```{r}
mod8_sd %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30.5, 25.5)))) %>% 
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()

# Rates of change
# Between 25.5 and 26.5
mod8_sd %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(26.5, 25.5))), type = "response") %>% 
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()

# Between 29.5 and 30.5
mod8_sd %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30.5, 29.5))), type = "response") %>% 
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()


```



# Save

```{r}
ggsave(mod8_sd_plot, file = "output/mod8_sd_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```


# Depth variability Pt II - Depth diff - don't really need this as the SD tells more of the story

# Fit candidate models
```{r}
mod8c <- glm(daily_depth_diff ~ SST, data = shark_daily_td)
mod8d <- gam(daily_depth_diff ~ s(SST, bs = "cs", k = 3), data = shark_daily_td)
```

## Validation and Information criteria
```{r}
# Dharma resids
mod8c %>% simulateResiduals(plot = TRUE) # 
mod8d %>% simulateResiduals(plot = TRUE) # 
# AIC
AIC(mod8c, mod8d)
performance::r2(mod8d)
```


These are basically the same but the GAM is slightly more complicated (obviously, with the basis terms etc). Check the summaries:
```{r}
mod8c |> summary()
mod8d |> summary()
```


The GAM edf is pretty low (1.6) but is significant -> means there is some small amount of wiggliness, but perhaps not enough to justify a GAM over a GLM.
We can plot to try to make more sense:

# The GLM (on the depth diff)
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod8c %>% 
  emmeans(~SST, at = SSTgrid) %>% 
  regrid() %>% 
  as_tibble() %>% 
    ggplot(aes(y = emmean, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_depth_diff, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Daily depth variation (m)") + 
    xlab("SST (°C)") +
    theme_bjc() +
    theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9))
mod8d %>% 
  emmeans(~SST, at = SSTgrid) %>% 
  regrid() %>% 
  as_tibble() %>% 
    ggplot(aes(y = emmean, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_depth_diff, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Daily depth variation (m)") + 
    xlab("SST (°C)") +
    theme_bjc() +
    theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9))


```

They both are telling the same story, but the GLM more parsiminious (less df) and more interpretable in this case (we don't really need a GAM here to tell us about changes in trends, inflection points etc)

```{r}
mod8_diff <- mod8c
```


# Planned contrast 26 - 30 ° C
```{r}

mod8_diff %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30, 26)))) %>% 
  #regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble()

```

So at 26°C SST variation in mean daily depth is ~70m. At 30°C it is 125m!! The diff is 55m 95%CI [50-62]


```{r}
# ENDS
```




