---
title: "Modelling depth use over time using GAM"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_width: 161/25.4
    fig_height: 100/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("code/packages.R")
#rm(list=ls()) # Clear out environment if necessary
```

# Load data 
## Using the shark_daily_grouped.Rmd
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
sourceRmd('code/01_eda/shark_daily_grouped.Rmd')
```

# Check df
```{r}
shark_daily_td
```


#  GRAPHICAL OVERVIEW #

## Daily depths (just the ggplot smooth)
```{r fig.width = 170/25.4, fig.height = 70/25.4}
shark_daily_depth_plot
```

# We need to formally model these trends with purpose-built GAMs

Notes:
- Likely going to use the same error distributions as the grouped glms but will check first
- Setting basis dimension (k), per months of the year - the sort of scale of trend we are interested in
- As we have a "cyclical" predictor variable (i.e. date) we'll need to insert a dummy variable (study_day) and then recode back after model
- For the same reason we'll set knots to "bridge" the gap between 31 Dec and 1 Jan and provide unique estimates at those points (not that important but doing it anyway)

# Dummy code study day to the main tbl
```{r}
shark_daily_td <- 
  shark_daily_td %>% 
  filter(date < as.Date("2022-10-31")) %>% 
  filter(date > as.Date("2021-10-30")) %>% 
  mutate(study_day = seq(1,365))
```


#### 1. Max depth ####

# Fit model - Gaussian
```{r}
maxdepth_timeGAM <- gam(daily_max_depth ~ s(study_day, bs = 'cc', k = 12), data = shark_daily_td, knots = list(study_day = c(0.5, 365.5)))
#maxdepth_timeGAMg <- gam(daily_max_depth ~ s(study_day, bs = 'cc', k = 12), data = shark_daily_td, family = Gamma, knots = list(study_day = c(0.5, 365.5)))
#maxdepth_timeGAMtw <- gam(daily_max_depth ~ s(study_day, bs = 'cc', k = 12), data = shark_daily_td, family = tw, knots = list(study_day = c(0.5, 365.5)))
#AIC(maxdepth_timeGAM, maxdepth_timeGAMg, maxdepth_timeGAMtw)
```

# Validation
```{r}
simulateResiduals(maxdepth_timeGAM) %>% plot()
gam.check(maxdepth_timeGAM)
```

## Results
```{r}
#maxdepth_timeGAM %>% tidy(conf.int=TRUE)) # Not that useful
maxdepth_timeGAM %>% summary()
```

# Make grid to use for all depth models
```{r fig.height = 56/25.4, fig.width = 56/25.4}
gam_grid <- shark_daily_td %>% 
  with(list(study_day = seq_range(study_day, n = 365)))
```

# New data for max depth 
```{r fig.height = 56/25.4, fig.width = 56/25.4}
maxdepthgam_newdata <- maxdepth_timeGAM %>% 
  emmeans(~study_day, at = gam_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(date = as.Date(study_day, origin = "2021-10-30")) # Add back in date
```

# Check output figure
```{r fig.width = 170/25.4, fig.height = 52/25.4}
    ggplot() +
    geom_line(data = shark_daily_td, aes(x = date, y = daily_max_depth)) +
    geom_ribbon(data = maxdepthgam_newdata, aes(y = emmean, x = date, ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.8) +
    geom_line(data = maxdepthgam_newdata, aes(y = emmean, x = date)) +
    ylab("Depth (m)") + 
    xlab("Day of year") +
    scale_y_reverse() +
    theme_minimal() +
    theme(plot.background = element_rect(colour = "black", linewidth = 0.5),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 9),
        legend.position = "none") 
```


# Identify peaks and troughs in maximum depths used:
```{r}
maxdepthgam_newdata %>% 
  slice_max(emmean, n = 7) # Max daily max depths of ~ 130m around mid-December

maxdepthgam_newdata %>% 
  slice_min(emmean, n = 7) # Minimum daily max depths of ~ 70m in late-August
```


#### 2. Mean depth ####

# Fit model - Gamma
```{r}
#meandepth_timeGAM <- gam(daily_mean_depth ~ s(study_day, bs = 'cc',  k = 12), data = shark_daily_td,  knots = list(study_day = c(0.5, 365.5)))
#meandepth_timeGAMg <- gam(daily_mean_depth ~ s(study_day, bs = 'cc',  k = 12), data = shark_daily_td, family = Gamma, knots = list(study_day = c(0.5, 365.5)))
#meandepth_timeGAMtw <- gam(daily_mean_depth ~ s(study_day, bs = 'cc',  k = 12), data = shark_daily_td, family = tw, knots = list(study_day = c(0.5, 365.5)))
#AIC(meandepth_timeGAM, meandepth_timeGAMg, meandepth_timeGAMtw)
meandepth_timeGAM <- gam(daily_mean_depth ~ s(study_day, bs = 'cc',  k = 12), data = shark_daily_td, family = Gamma, knots = list(study_day = c(0.5, 365.5)))
```

# Valdation
```{r}
simulateResiduals(meandepth_timeGAM) %>% plot()
gam.check(meandepth_timeGAM)
```


```{r}
## Results
meandepth_timeGAM %>% tidy(conf.int=TRUE)
meandepth_timeGAM %>% summary()
```


```{r fig.height = 56/25.4, fig.width = 56/25.4}
meandepthgam_newdata <- meandepth_timeGAM %>% 
  emmeans(~study_day, at = gam_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(date = as.Date(study_day, origin = "2021-10-30"))
```

# Figure for depth variation by time of day
```{r fig.width = 170/25.4, fig.height = 52/25.4}
    ggplot() +
    geom_point(data = shark_daily_td, aes(x = study_day, y = daily_mean_depth), pch = 21, alpha = 0.5) +
    geom_ribbon(data = meandepthgam_newdata, aes(y = response, x = study_day, ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.8) +
    geom_line(data = meandepthgam_newdata, aes(y = response, x = study_day)) +
    ylab("Depth (m)") + 
    xlab("Day of year") +
   # scale_x_continuous(limits = c(-1, 24), breaks = c(0, 6, 12, 18, 24)) +
    theme_minimal() +
    scale_y_reverse() +
    theme(plot.background = element_rect(colour = "black", linewidth = 0.5),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 9),
        legend.position = "none") 
```



# Identify peaks and troughs in maximum depths used:

```{r}
meandepthgam_newdata %>% 
  slice_max(response, n = 7) # Deepest mean depths of ~30 m in mid-January
meandepthgam_newdata %>% 
  slice_min(response, n = 7) # Shallowest mean depths of ~15m in late-Sept
```


#### 3. Minimum depth ####

# Fit model - negbinom
```{r}
mindepth_timeGAM <- gam(daily_min_depth ~ s(study_day, bs = 'cc',  k = 12), data = shark_daily_td, family = negbin(1), knots = list(study_day = c(0.5, 365.5)))
#mindepth_timeGAM2 <- gam(daily_min_depth ~ s(study_day, bs = 'cc',  k = 12), data = shark_daily_td, family = negbin(2), knots = list(study_day = c(0.5, 365.5)))
#AIC(mindepth_timeGAM1, mindepth_timeGAM2)
```

# Valdation
```{r}
simulateResiduals(mindepth_timeGAM) %>% plot()
gam.check(mindepth_timeGAM)
```


```{r}
## Results
#mindepth_timeGAM %>% tidy(conf.int=TRUE)
mindepth_timeGAM %>% summary()
```


```{r fig.height = 56/25.4, fig.width = 56/25.4}
mindepthgam_newdata <- 
  mindepth_timeGAM %>% 
  emmeans(~study_day, at = gam_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(date = as.Date(study_day, origin = "2021-10-30"))

```


# Figure for depth variation by time of day
```{r fig.width = 170/25.4, fig.height = 52/25.4}
    ggplot() +
    geom_line(data = shark_daily_td, aes(x = date, y = daily_min_depth)) +
    geom_ribbon(data = mindepthgam_newdata, aes(y = response, x = date, ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.8) +
    geom_line(data = mindepthgam_newdata, aes(y = response, x = date)) +
    ylab("Temp (°C)") + 
    xlab("Day of year") +
    scale_y_reverse() +
    theme_minimal() +
    theme(plot.background = element_rect(colour = "black", linewidth = 0.5),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 9),
        legend.position = "none") 
```

Not much point in presenting the results of this - just outliers pulling CI down but only to 3 meters or so! Won't even see it on the main figure.


## Redo the time series fig with the correct GAM data included
## Daily temps (all) vs SST
```{r fig.width = 170/25.4, fig.height = 70/25.4}
shark_daily_depths_plot <- 
  ggplot() +
  geom_line(data = shark_daily_td, aes(x = date, y = daily_min_depth), colour = "skyblue") +
  geom_ribbon(data = meandepthgam_newdata, aes(y = response, x = date, ymin=lower.CL, ymax=upper.CL), alpha=0.4, fill = "dodgerblue", colour = "NA") +
  geom_line(data = meandepthgam_newdata, aes(x = date, y = response), colour = "dodgerblue") +
  geom_line(data = shark_daily_td, aes(x = date, y = daily_mean_depth), colour = "dodgerblue") +
  geom_ribbon(data = maxdepthgam_newdata, aes(y = emmean, x = date, ymin=lower.CL, ymax=upper.CL), alpha=0.4, fill = "midnightblue", colour = "NA") +
  geom_line(data = shark_daily_td, aes(x = date, y = daily_max_depth), colour = "midnightblue") +
  geom_line(data = maxdepthgam_newdata, aes(y = emmean, x = date), colour = "midnightblue") +
  theme_minimal() +
  scale_y_reverse() +
  scale_x_date(date_breaks = "1 months", date_labels = '%b', expand = c(0.02, 0.02)) +
  theme(axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 9, colour = "black")) +
  xlab("Month of year (2021 - 2022)") +
  ylab("Depth (m below sea level)")

shark_daily_depths_plot
```

```{r}
#END
```

