---
title: "GAM for depth movement ~ TOD"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 3
    fig_width: 9
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
getwd()
source("code/packages.R")
#rm(list=ls()) # Clear out environment if necessary
```


# Load data
```{r}
getwd()
sourceRmd("code/01_eda/eda_tod.Rmd")
sourceRmd("code/00_wrangle/wrangling_SSTdata.Rmd")
```

# Initial checks
```{r}
shark_tp_data  %$% 
  summary(hour)

shark_tp_data %>% 
  ggplot() +
  geom_histogram(aes(x = Depth))
```

```{r}
depth_gam_data <- shark_tp_data %>% 
  filter(!is.na(Depth)) %>% 
  mutate(Season = fct_relevel(Season, c("Summer", "Autumn", "Winter", "Spring")))
```


```{r}
tod_gam <- gam(Depth ~ Season + s(hour, bs = "cs", by = Season), data = depth_gam_data, family = nb, method = "REML")
```



```{r}
#simulateResiduals(tod_gam) %>% plot()
gam.check(tod_gam)
```

Method: REML   Optimizer: outer newton
full convergence after 5 iterations.
Gradient range [6.200262e-05,0.0580288]
(score 1436359 & scale 1).
Hessian positive definite, eigenvalue range [4.151618,153972.2].
Model rank =  40 / 40 

Basis dimension (k) checking results. Low p-value (k-index<1) may
indicate that k is too low, especially if edf is close to k'.

                       k'  edf k-index p-value
s(hour):SeasonAutumn 9.00 8.91    0.96    0.58
s(hour):SeasonSpring 9.00 8.73    0.96    0.52
s(hour):SeasonSummer 9.00 8.89    0.96    0.55
s(hour):SeasonWinter 9.00 8.88    0.96    0.54
```{r}
## Results
(tod_gamresults <- tod_gam %>% tidy(conf.int=TRUE))
tod_gam %>% summary()
```




```{r fig.height = 56/25.4, fig.width = 56/25.4}
tod_gam_grid <- depth_gam_data %>% 
  with(list(Season = levels(Season),
            hour = seq_range(hour, n = 100)))

depgam_newdata <- tod_gam %>% 
  emmeans(~hour * Season, at = tod_gam_grid) %>% 
  regrid() %>% 
  as_tibble()
```


# Results (peaks) from depgam_newdata
```{r}
depgam_peaks <- 
  depgam_newdata %>% 
  ungroup() %>% 
  mutate(hour = floor(hour)) %>% 
  group_by(hour, Season, lower.CL, upper.CL) %>% 
  summarise(max = max(response)) %>% # Get multiple values of Depth (response) per hour as we've "floored" the time stamp
  group_by(Season) %>% 
  slice_max(max) %>% 
  mutate(`±CI95%` = upper.CL-max)

depgam_lows <- 
  depgam_newdata %>% 
  ungroup() %>% 
  mutate(hour = floor(hour)) %>% 
  group_by(hour, Season, lower.CL, upper.CL) %>% 
  summarise(min = min(response)) %>% # Get multiple values of Depth (response) per hour as we've "floored" the time stamp
  group_by(Season) %>% 
  slice_min(min) %>% 
  mutate(`±CI95%` = upper.CL-min)


 
```

```{r}
depgam_peaks
depgam_lows
```




# Figure for depth variation by time of day
```{r fig.width = 170/25.4, fig.height = 52/25.4}
depgam_plot <- 
    ggplot() +
    annotate("rect", xmin = 0, xmax = 6, ymin = 0, ymax = 80, fill = "grey50", alpha = 0.5) +
    annotate("rect", xmin = 18, xmax = 24, ymin = 0, ymax = 80, fill = "grey50", alpha = 0.5) +
    geom_errorbar(data = hourly_shark_data, aes(ymin = mean_depth-sd_depth, ymax = mean_depth+sd_depth, x = hour), width = 1.5, colour = "grey40", linewidth = 0.2) +
    geom_point(data = hourly_shark_data, aes(x = hour, y = mean_depth, fill = Season), pch = 21, alpha = 0.5) +
    geom_ribbon(data = depgam_newdata, aes(y = response, x = hour, ymin=lower.CL, ymax=upper.CL, fill = Season), color=NA, alpha=0.8) +
    geom_line(data = depgam_newdata, aes(y = response, x = hour)) +
    ylab("Depth (m)") + 
    xlab("Time of day") +
    scale_fill_manual(values = c("dodgerblue","red", "orange", "darkblue")) +
    scale_y_reverse() +
    scale_x_continuous(limits = c(-1, 24), breaks = c(0, 6, 12, 18, 24)) +
    theme_minimal() +
    theme(plot.background = element_rect(colour = "black", linewidth = 0.5),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 9),
        legend.position = "none") +
    facet_wrap(~ Season, ncol = 4)

depgam_plot

ggsave(depgam_plot, file = "output/dep_gam_only_fig.pdf", width = 170, height = 60, device = cairo_pdf, units = "mm", dpi = 300)

```

# Variation in depth use by hour ~ season (uses SD of mean depth)
```{r fig.height=50/25.4, fig.width=160/25.4}
hr_sddepth_plot <- hourly_shark_data %>% 
  ggplot() +
  annotate("rect", xmin = 0, xmax = 6, ymin = 0, ymax = 35, fill = "grey50", alpha = 0.5) +
  annotate("rect", xmin = 18, xmax = 24, ymin = 0, ymax = 35, fill = "grey50", alpha = 0.5) +
  geom_line(aes(x = hour, y = sd_depth), colour = "grey20") +
  geom_point(aes(x = hour, y = sd_depth, fill = Season), pch = 21, alpha = 0.5) +
  scale_fill_manual(values = c("dodgerblue","red", "orange", "darkblue")) +
  #scale_y_reverse() +
  theme_minimal() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) +
  facet_wrap(~ Season, ncol = 4) +
  scale_x_continuous(limits = c(-1, 24), breaks = c(0, 6, 12, 18, 24)) +
  theme(legend.position = "none",
        strip.text = element_blank()) +
  ylab("Std. dev. depth (m)") + 
  xlab("Time of day")

hr_sddepth_plot
```

# Combine in Cowplot
```{r}
tod_dep_plot <- cowplot::plot_grid(depgam_plot, hr_sddepth_plot, nrow = 2, labels = c('A', 'B'), label_size = 9, align = "v") +
  theme(plot.background = element_rect(color = "black"))

```

```{r fig.height = 140/25.4, fig.width = 140/25.4}
#depgam_plot
#ggsave(tod_dep_plot, file = "output/dep_gam_sd_fig.pdf", width = 170, height = 100, device = cairo_pdf, units = "mm", dpi = 300)
```



### ALTERNATIVE PLOT ###


```{r}
hr_dep_sst_data <- 
depth_gam_data %>% 
  left_join(SST) %>% 
  mutate(hour = as.numeric(hour))

hr_dep_sst_data %$% 
  summary(Season)
depgam_newdata  %$% 
  summary(Season)
```

## Need to load in sun and moon files
```{r}
library(png)
library(grid)
sun <- readPNG("data/sun.png") %>% grid::rasterGrob(interpolate = TRUE)
moon <- readPNG("data/moon.png") %>% grid::rasterGrob(interpolate = TRUE)
```

# Now plot
```{r fig.height=100/25.4, fig.width=120/25.4}
hr_dep_sst_plot <- 
  ggplot() +
  annotate("rect", xmin = 0, xmax = 24, ymin = 0, ymax = 150, fill = "grey70", alpha = 0.4) + # Boxes to indicate night vs day. Need to start at -0.3 to reach underneath jittered pts
  annotate("rect", xmin = 0, xmax = 6, ymin = 0, ymax = 150, fill = "grey40", alpha = 0.3) + # Boxes to indicate night vs day 
  annotate("rect", xmin = 18, xmax = 24, ymin = 0, ymax = 150, fill = "grey40", alpha = 0.3) + # Boxes to indicate night vs day 
  geom_point(data = hr_dep_sst_data, aes(x = hour, y = Depth, colour = SST), pch = 20, alpha = 0.8, size = 0.1, position = "jitter") +
 # geom_point(data = hr_dep_sst_data, aes(x = hour, y = Depth, colour = SST), fill = "transparent", pch = 22, alpha = 0.7, position = "jitter") +
  scale_fill_gradient2(low = "darkblue", mid = "transparent", high = "red2", midpoint = 28)  +
  scale_colour_gradient2(low = "darkblue", mid = "transparent", high = "red2", midpoint = 28)  + 
  #scale_fill_gradient(low = "blue", high = "red2", aesthetics = c("colour", "fill"))  +
  #annotate("rect", xmin = -0.5, xmax = -0.1, ymin = -0.3, ymax = 150,  fill = "white") +
  geom_errorbar(data = hourly_shark_data, aes(ymin = mean_depth-sd_depth, ymax = mean_depth+sd_depth, x = hour), width = 0.5, colour = "black", linewidth = 0.2) +
  geom_point(data = hourly_shark_data, aes(x = hour, y = mean_depth), colour = "black", fill = "white", pch = 21, alpha = 0.9) +
  #geom_ribbon(data = depgam_newdata, aes(y = response, x = hour, ymin=lower.CL, ymax=upper.CL), colour = "grey40", fill = "grey40", alpha = 0.5) +
  geom_line(data = depgam_newdata, aes(y = response, x = hour), colour = "black", linewidth = 0.5) +
  scale_y_reverse(limits = c(150, -0.3), expand = c(0,0)) +
  scale_x_continuous(limits = c(-0.3, 24), breaks = c(0, 6, 12, 18, 24)) +
  annotation_custom(moon, xmin = 1.5, xmax = 4.5, ymax = -120) +
  annotation_custom(moon, xmin = 19.5, xmax = 22.5, ymax = -120) +
  annotation_custom(sun, xmin = 10, xmax = 14, ymax = -120) +
  theme_minimal() +
  theme(#panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5),    # Box around each plot
        plot.background = element_rect(colour = "black", linewidth = 0.5),             # Box around whole plot
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) +
  theme(legend.title = element_text(size = 9), 
        #legend.position = "none", 
        legend.text = element_text(size = 9)) +
  facet_wrap(~ Season, ncol = 2, nrow = 2) +
  ylab("Depth (m)") + 
  xlab("Time of day")

hr_dep_sst_plot
```




```{r}
#ggsave(hr_dep_sst_plot, file = "output/hr_dep_sst_plot_final.pdf", width = 170, height = 54, device = cairo_pdf, units = "mm", dpi = 300)
ggsave(hr_dep_sst_plot, file = "output/hr_dep_sst_plot_final_2col.pdf", device = cairo_pdf, width = 120, height = 100, units = "mm", dpi = 300)
ggsave(hr_dep_sst_plot, file = "output/hr_dep_sst_plot_final_2col.jpg", device = "jpeg", width = 120, height = 100, units = "mm", dpi = 300)
ggsave(hr_dep_sst_plot, file = "output/hr_dep_sst_plot_final_2col.png", device = png, width = 120, height = 100, units = "mm", dpi = 300)
```





