---
title: "Modelling internal body temperatures over time using GAM"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_width: 161/25.4
    fig_height: 100/25.4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("code/packages.R")
#rm(list=ls()) # Clear out environment if necessary
```

# Load data 
## Using the shark_daily_grouped.Rmd
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=6}
sourceRmd('code/01_eda/shark_daily_grouped.Rmd')
```

# Check df
```{r}
shark_daily_td
```


#  GRAPHICAL OVERVIEW #

## Daily temps (all) vs SST
```{r fig.width = 170/25.4, fig.height = 70/25.4}
shark_daily_temps_plot 
```

# We need to formally model these trends with purpose-built GAMs

Notes:
- Likely going to use the same error distributions as the grouped glms but will check first
- Setting basis dimension (k), per months of the year - the sort of scale of trend we are interested in
- As we have a "cyclical" predictor variable (i.e. date) we'll need to insert a dummy variable (study_day) and then recode back after model
- For the same reason we'll set knots to "bridge" the gap between 31 Dec and 1 Jan and provide unique estimates at those points (not that important but doing it anyway)

# Dummy code study day to the main tbl
```{r}
shark_daily_td <- 
  shark_daily_td %>% 
  filter(date < as.Date("2022-10-31")) %>% 
  filter(date > as.Date("2021-10-30")) %>% 
  mutate(study_day = seq(1,365))
```



# Apply GAMs

#### 1. Minimum Tb ####

# Fit model
```{r}
mintemp_timeGAM <- gam(daily_min_temp ~ s(study_day, bs = 'cc', k = 12), data = shark_daily_td, knots = list(doy = c(0.5, 365.5)))
```

# Validation
```{r}
simulateResiduals(mintemp_timeGAM) %>% plot()
gam.check(mintemp_timeGAM)
```

## Results
```{r}
#mintemp_timeGAM %>% tidy(conf.int=TRUE))
mintemp_timeGAM %>% summary()
```

# Make grid to use for all depth models
```{r fig.height = 56/25.4, fig.width = 56/25.4}
gam_grid <- shark_daily_td %>% 
  with(list(study_day = seq_range(study_day, n = 365)))
```

# New date for min Tb
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mintempgam_newdata <- mintemp_timeGAM %>% 
  emmeans(~study_day, at = gam_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(date = as.Date(study_day, origin = "2021-10-30")) # Add back in date
```


# Figure for depth variation by time of day
```{r fig.width = 170/25.4, fig.height = 52/25.4}
#time_temp_gam_plot <- 
    ggplot() +
    geom_line(data = shark_daily_td, aes(x = date, y = daily_min_temp)) +
    geom_ribbon(data = mintempgam_newdata, aes(y = emmean, x = date, ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.8) +
    geom_line(data = mintempgam_newdata, aes(y = emmean, x = date)) +
    ylab("Temp (°C)") + 
    xlab("Day of year") +
   # scale_x_continuous(limits = c(-1, 24), breaks = c(0, 6, 12, 18, 24)) +
    theme_minimal() +
    theme(plot.background = element_rect(colour = "black", linewidth = 0.5),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 9),
        legend.position = "none") 
```




```{r}
mintempgam_newdata %>% 
  slice_max(emmean, n = 1) # Max daily min Tb 26.0°C

mintempgam_newdata %>% 
  slice_min(emmean, n = 1) # Min daily min Tb 24.9
```


#### 2. Mean Tb ####

# Fit model
```{r}
meantemp_timeGAM <- gam(daily_mean_temp ~ s(study_day, bs = 'cc'), data = shark_daily_td, knots = list(study_day = c(0.5, 365.5)))
```

# Valdation
```{r}
simulateResiduals(meantemp_timeGAM) %>% plot()
gam.check(meantemp_timeGAM)
```


```{r}
## Results
(gamresults <- meantemp_timeGAM %>% tidy(conf.int=TRUE))
meantemp_timeGAM %>% summary()
```


```{r fig.height = 56/25.4, fig.width = 56/25.4}
meantempgam_newdata <- meantemp_timeGAM %>% 
  emmeans(~study_day, at = gam_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(date = as.Date(study_day, origin = "2021-10-30"))
```



# Figure for depth variation by time of day
```{r fig.width = 170/25.4, fig.height = 52/25.4}
time_temp_gam_plot <- 
    ggplot() +
    geom_point(data = shark_daily_td, aes(x = study_day, y = daily_mean_temp), pch = 21, alpha = 0.5) +
    geom_ribbon(data = meantempgam_newdata, aes(y = emmean, x = study_day, ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.8) +
    geom_line(data = meantempgam_newdata, aes(y = emmean, x = study_day)) +
    ylab("Temp (°C)") + 
    xlab("Day of year") +
   # scale_x_continuous(limits = c(-1, 24), breaks = c(0, 6, 12, 18, 24)) +
    theme_minimal() +
    theme(plot.background = element_rect(colour = "black", linewidth = 0.5),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 9),
        legend.position = "none") 

time_temp_gam_plot
```




```{r}
meantempgam_newdata %>% 
  slice_max(emmean, n = 10)# Max mean temperature 28.85°C is almost exactly the same as annual MMM (28.69°C)

```


#### 3. Max Tb ####

# Fit model
```{r}
maxtemp_timeGAM <- gam(daily_max_temp ~ s(study_day, bs = 'cc'), data = shark_daily_td, knots = list(study_day = c(0.5, 365.5)))
```

# Valdation
```{r}
simulateResiduals(maxtemp_timeGAM) %>% plot()
gam.check(maxtemp_timeGAM)
```


```{r}
## Results
(gamresults <- maxtemp_timeGAM %>% tidy(conf.int=TRUE))
maxtemp_timeGAM %>% summary()
```


```{r fig.height = 56/25.4, fig.width = 56/25.4}
maxtempgam_newdata <- 
  maxtemp_timeGAM %>% 
  emmeans(~study_day, at = gam_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(date = as.Date(study_day, origin = "2021-10-30"))

```



# Figure for depth variation by time of day
```{r fig.width = 170/25.4, fig.height = 52/25.4}
time_maxTb_plot <- 
    ggplot() +
    geom_point(data = shark_daily_td, aes(x = study_day, y = daily_max_temp), pch = 21, alpha = 0.5) +
    geom_ribbon(data = maxtempgam_newdata, aes(y = emmean, x = study_day, ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.8) +
    geom_line(data = maxtempgam_newdata, aes(y = emmean, x = study_day)) +
    ylab("Temp (°C)") + 
    xlab("Day of year") +
   # scale_x_continuous(limits = c(-1, 24), breaks = c(0, 6, 12, 18, 24)) +
    theme_minimal() +
    theme(plot.background = element_rect(colour = "black", linewidth = 0.5),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 9),
        legend.position = "none") 

time_maxTb_plot
```


# Extract key findings
```{r}
# Model points towards sustained high temps immediately before TC Seth
maxtempgam_newdata %>% 
  slice_max(emmean, n = 12) %>%  # Max modeled max temperature 30°C
  arrange(date)

# Concurs with raw data:
shark_daily_td %>% 
  slice_max(SST, n = 5)  # Max modeled max temperature 30°C



```

## Redo the time series fig with the correct GAM data included
## Daily temps (all) vs SST
```{r fig.width = 170/25.4, fig.height = 70/25.4}
shark_daily_temps_plot <- 
  ggplot() +
  geom_line(data = shark_daily_td, aes(x = date, y = SST), colour = "grey15", lwd = 0.6, lty = 3313) +
  geom_ribbon(data = maxtempgam_newdata, aes(y = emmean, x = date, ymin=lower.CL, ymax=upper.CL), alpha=0.4, fill = "red", colour = "NA") +
  geom_line(data = maxtempgam_newdata, aes(y = emmean, x = date), colour = "red") +
  geom_line(data = shark_daily_td, aes(x = date, y = daily_max_temp), colour = "red") +
  geom_ribbon(data = mintempgam_newdata, aes(y = emmean, x = date, ymin=lower.CL, ymax=upper.CL), alpha=0.4, fill = "midnightblue", colour = "NA") +
  geom_line(data = shark_daily_td, aes(x = date, y = daily_min_temp), colour = "midnightblue") +
  geom_line(data = mintempgam_newdata, aes(y = emmean, x = date), colour = "midnightblue") +
  geom_ribbon(data = meantempgam_newdata, aes(y = emmean, x = date, ymin=lower.CL, ymax=upper.CL), alpha=0.4, fill = "grey40", colour = "NA") +
  geom_line(data = shark_daily_td, aes(x = date, y = daily_mean_temp), colour = "grey40") +
  geom_line(data = meantempgam_newdata, aes(y = emmean, x = date), colour = "grey40") +
  scale_y_continuous(breaks = c(24,25,26,27,28,29,30,31)) +
  scale_x_date(date_breaks = "1 months", date_labels = '%b', expand = c(0.02, 0.02)) +
  theme_minimal() +
  theme(axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 9, colour = "black")) +
  xlab("Month of year (2021 - 2022)") +
  ylab(expression(paste(italic("T"),""[b]* " (°C)")))

shark_daily_temps_plot
```

```{r}
#END
```

