---
title: "Shark daily internal temperature ~ SST models"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
source("code/packages.R")
#rm(list=ls()) # Clear out environment if necessary
```

# Load data
```{r}
load(file = "data/Rdata/shark_daily_td.Rdata")
shark_daily_td
```


# We need models for 

1. Max daily temp ~ SST    - Gaussian
2. Mean daily temp ~ SST   - Gaussian
3. Min daily temp ~ SST    - 
4. Daily temp SD ~ SST

```{r}
## Check distributions so we know how to model
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_max_temp))
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_mean_temp))
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_min_temp))
```

All look approximately normal

##### Model 1: Daily max temp ~ SST #####

## Model formula - gaussian best fit and AIC score
```{r}
mod1 <- glm(daily_max_temp ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod1 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod1 %>% performance::check_outliers()
mod1 %>% performance::r2()
```

All look approximately normal and R2: 0.953


## Results
Almost 1:1 relationship between SST and max internal body temperature 
```{r}
(mod1results <- mod1 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod1.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata1 <- mod1 %>% 
  emmeans(~SST, at = mod1.grid) %>% 
  regrid() %>% 
  as_tibble()

mod1_plot <- 
  ggplot(newdata1, aes(y = emmean, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_max_temp, x = SST), fill = "red", col = "black", alpha = 0.4, pch = 21) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
  geom_line(colour = "black") +
 # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  #ylab("Daily max temp (°C)") + 
  ylab(expression(paste("Daily max", italic("T"),""[b]* " (°C)"))) + 
  xlab("SST (°C)") +
  scale_y_continuous(limits = c(25.4, 30.9), n.breaks = 6) +
  scale_x_continuous(limits = c(25.4, 30.9), n.breaks = 6) +
  theme_minimal() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 9, colour = "black"))

mod1_plot

ggsave(mod1_plot, file = "output/mod1_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```

## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
#(daily_max_temp_model_plot <- ggarrange(shark_daily_maxtemp_plot, mod1_plot, ncol = 2, widths = c(2,1)))
#ggsave(daily_max_temp_model_plot, file = "output/daily_max_temp_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)
```


##### Model 2a: Mean daily temp ~ SST #####

## Model formula - again gaussian best fit
```{r}
mod2a <- glm(daily_mean_temp ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod2a %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod2a %>% performance::check_outliers()
mod2a %>% performance::r2()
```
All look approximately normal

## Results
```{r}
(mod2aresults <- mod2a %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod2a.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata2a <- mod2a %>% 
  emmeans(~SST, at = mod2a.grid) %>% 
  regrid() %>% 
  as_tibble()

mod2a_plot <- 
  ggplot(newdata2a, aes(y = emmean, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_mean_temp, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
  geom_line() +
  # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  #ylab("Daily mean temp (°C)") + 
  ylab(expression(paste("Daily mean", italic("T"),""[b]* " (°C)"))) + 
  xlab("SST (°C)") +
  scale_y_continuous(limits = c(25.4, 30.9), n.breaks = 6) +
  scale_x_continuous(limits = c(25.4, 30.9), n.breaks = 6) +
  theme_minimal() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 9, colour = "black"))

mod2a_plot

ggsave(mod2a_plot, file = "output/mod2a_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```

## Add to time-series data if required
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
#(daily_mean_temp_model_plot <- ggarrange(shark_daily_meantemp_plot, mod2a_plot, ncol = 2, widths = c(2,1)))
#ggsave(daily_mean_temp_model_plot, file = "output/daily_mean_temp_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)
```



##### Model 2b: Mean daily temp SD ~ SST #####

## Model formula
We know from EDA that we'll need a GAM for this. But what error family?

```{r}
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_temp_sd))
```
Looks right skewed. Maybe bimodal?

```{r}
mod2bnorm <- gam(daily_temp_sd ~ s(SST, bs = "cs"), data = shark_daily_td, method = "REML")
mod2btw <- gam(daily_temp_sd ~ s(SST, bs = "cs"), data = shark_daily_td, family = tw, method = "REML")
AIC(mod2bnorm, mod2btw)
mod2b <- mod2btw
```

## Model validation
```{r}
# Dharma resids
mod2b %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod2b %>% performance::check_outliers()
mod2b %>% gam.check()
mod2b %>% performance::r2()
```

## Results
```{r}
(mod2bresults <- mod2b %>% tidy(conf.int=TRUE))
mod2b %>% summary()
mod2bnorm %>% summary()
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod2b.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata2b <- mod2b %>% 
  emmeans(~SST, at = mod2b.grid) %>% 
  regrid() %>% 
  as_tibble()

mod2b_plot <- 
    ggplot(newdata2b, aes(y = response, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_temp_sd, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
   # ylab("Std. dev. daily temp (°C)") + 
    ylab(expression(paste("S.D. daily mean", italic("T"),""[b]* " (°C)"))) + 
    xlab("SST (°C)") +
    theme_minimal() +
    theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 9, colour = "black"))


mod2b_plot

ggsave(mod2b_plot, file = "output/mod2b_plot.pdf", width = 62, height = 62, device = cairo_pdf, units = "mm", dpi = 300)
```
# Temp GAM results - planned contrasts 
```{r}
mod2b.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

mod2b %>% 
  emmeans(~SST, at = shark_daily_td %>% 
            with(list(SST = c(30, 26))), type = "response") %>% 
  regrid() %>% 
  #pairs() %>% 
  confint() %>% 
  as_tibble()

```


##### Model 3: Min Daily Temp ~ SST #####

## Model formula - gaussian best fit
```{r}
mod3 <- glm(daily_min_temp ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod3 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod3 %>% performance::check_outliers()
mod3 %>% performance::r2()
```


## Results
```{r}
(mod3results <- mod3 %>% tidy(conf.int=TRUE))
mod3 %>% summary()
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod3.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata3 <- mod3 %>% 
  emmeans(~SST, at = mod3.grid) %>% 
  regrid() %>% 
  as_tibble()

mod3_plot <- 
  ggplot(newdata3, aes(y = emmean, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_min_temp, x = SST), fill = "midnightblue", col = "black", alpha = 0.4, pch = 21) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
  geom_line() +
  theme_minimal() +
 # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
 # ylab("Daily min temp (°C)") + 
  ylab(expression(paste("Daily min", italic("T"),""[b]* " (°C)"))) + 
  xlab("SST (°C)") +
  #scale_y_continuous(limits = c(25, 30.5), n.breaks = 6) +
  #scale_x_continuous(limits = c(25, 30.5), n.breaks = 6) +
  theme_minimal() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 9, colour = "black"))

mod3_plot

ggsave(mod3_plot, file = "output/mod3_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```




## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
(daily_min_temp_model_plot <- ggarrange(shark_daily_mintemp_plot, mod3_plot, ncol = 2, widths = c(2,1)))
ggsave(daily_min_temp_model_plot, file = "output/daily_min_temp_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)

```




```{r}
# ENDS
```

