---
title: "Shark daily temperature ~ depth models"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
#rm(list=ls()) # Clear out environment if necessary
```


```{r include=FALSE}
source("code/packages.R")
source("code/extras.R")
```

# Load data
```{r}
load(file = "data/Rdata/shark_daily_td.Rdata")
shark_daily_td
```


# We need models for 

1. Max daily temp ~ SST
2. Mean daily temp ~ SST
3. Min daily temp ~ SST
4. Max daily depth ~ SST

```{r}
## Check distributions so we know how to model
## Max temp
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_max_temp))
## Mean temp
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_mean_temp))
## Min temp
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_min_temp))
## SD temp
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_temp_sd))
```

All look approximately normal

##### Model 1: Daily max temp ~ SST #####

## Model formula
```{r}
mod1 <- glm(daily_max_temp ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod1 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod1 %>% performance::check_outliers()
mod1 %>% performance::r2()
```

All look approximately normal and R2: 0.953


## Results
Almost 1:1 relationship between SST and max internal body temperature 
```{r}
(mod1results <- mod1 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height=80/25.4, fig.width=80/25.4}
mod1.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata1 <- mod1 %>% 
  emmeans(~SST, at = mod1.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod1_plot <- 
  ggplot(newdata1, aes(y = emmean, x = SST)) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
  geom_line() +
  geom_point(data=shark_daily_td,  aes(y=daily_max_temp, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
  theme_bjc() +
 # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  ylab("Daily max shark temp (°C)") + 
  xlab("\nSST (°C)"))
```


```{r fig.height=80/25.4, fig.width=80/25.4}
ggsave(mod1_plot, file = "output/mod1_plot.pdf", 
       width = 100, height = 100, units = "mm", dpi = 300)
```

## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
(daily_max_temp_model_plot <- ggarrange(shark_daily_maxtemp_plot, mod1_plot, ncol = 2, widths = c(2,1)))
ggsave(daily_max_temp_model_plot, file = "output/daily_max_temp_model_plot.pdf", 
       width = 180, height = 60, units = "mm", dpi = 300)
```


##### Model 2a: Mean daily temp ~ SST #####

## Model formula
Will try glm and simple gam
```{r}
mod2ai <- glm(daily_mean_temp ~ SST, data = shark_daily_td, family = gaussian)
mod2aii <- gam(daily_mean_temp ~ s(SST, bs = "cs", k =3), data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod2ai %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod2ai %>% performance::check_outliers()
mod2ai %>% performance::r2()

mod2aii %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod2aii %>% performance::check_outliers()
mod2aii %>% performance::r2()

AIC(mod2ai, mod2aii)
```



All look approximately normal


## Results
```{r}
(mod2aresults <- mod2a %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height=80/25.4, fig.width=80/25.4}
mod2a.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata2a <- mod2ai %>% 
  emmeans(~SST, at = mod2a.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod2ai_plot <- 
    ggplot(newdata2a, aes(y = emmean, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_mean_temp, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    theme_bjc() +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Daily mean temp (°C)") + 
    xlab("\nSST (°C)"))
```

# Save
```{r fig.height=80/25.4, fig.width=80/25.4}
ggsave(mod2a_plot, file = "output/mod2a_plot.pdf", 
       width = 100, height = 100, units = "mm", dpi = 300)
```

## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
(daily_mean_temp_model_plot <- ggarrange(shark_daily_meantemp_plot, mod2a_plot, ncol = 2, widths = c(2,1)))
ggsave(daily_mean_temp_model_plot, file = "output/daily_mean_temp_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)

```




##### Model 2b: Mean daily temp SD ~ SST #####

## Model formula
Need a GAM


```{r}
mod2bi <- gam(daily_temp_sd ~ s(SST, bs = "cs", k = 3), family = tw(), data = shark_daily_td)
mod2bii <- gam(daily_temp_sd ~ s(SST, bs = "cs", k = 3), family = Gamma(), data = shark_daily_td)
```

## Model validation
```{r}
# Dharma resids
mod2bi %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod2bi %>% performance::check_outliers()
mod2bi %>% performance::r2()

mod2bii %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod2bii %>% performance::check_outliers()
mod2bii %>% performance::r2()

AIC(mod2bi, mod2bii)

mod2b <- mod2bii
```

## Results
```{r}
(mod2bresults <- mod2b %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height=80/25.4, fig.width=80/25.4}
mod2b.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata2b <- mod2b %>% 
  emmeans(~SST, at = mod2b.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod2b_plot <- 
    ggplot(newdata2b, aes(y = response, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_temp_sd, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    theme_bjc() +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Std deviation in\ndaily temps (°C)") + 
    xlab("\nSST (°C)"))
```


```{r fig}
ggsave(mod2b_plot, file = "output/mod2b_plot.pdf", 
       width = 80, height = 80, units = "mm", dpi = 300)
```



##### Model 2b: Delta between Mean and Max daily temps ~ SST #####
### NOT GOING TO USE THIS - SD SHOULD BE SUFFICIENT


## Model formula
Try with GAM
```{r}
mod2c <- gam(temp_diff ~ s(SST, bs = "cs", k =3), data = shark_daily_td %>% 
  mutate(temp_diff = daily_max_temp - daily_min_temp))
```

## Model validation
```{r}
# Dharma resids
mod2c %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod2c %>% performance::check_outliers()
mod2c %>% performance::r2()
```

## Results
```{r}
(mod2cresults <- mod2c %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height=80/25.4, fig.width=80/25.4}
mod2c.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata2c <- mod2c %>% 
  emmeans(~SST, at = mod2c.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod2c_plot <- 
    ggplot(newdata2c, aes(y = emmean, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td %>% mutate(temp_diff = daily_max_temp - daily_min_temp),  aes(y=temp_diff, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    theme_bjc() +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Diff between min and\nmax daily temps (°C)") + 
    xlab("\nSST (°C)"))
ggsave(mod2c_plot, file = "output/mod2c_plot.pdf", width = 100, height = 100, units = "mm", dpi = 300)
```



##### Model 3: Min Daily Temp ~ SST #####

## Model formula
```{r}
mod3i <- glm(daily_min_temp ~ SST, data = shark_daily_td, family = gaussian)
mod3ii <- gam(daily_min_temp ~ s(SST, bs = "cs", k = 3), data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod3i %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod3i %>% performance::check_outliers()
mod3i %>% performance::r2()

mod3ii %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod3ii %>% performance::check_outliers()
mod3ii %>% performance::r2()

AIC(mod3i, mod3ii)

mod3 <- mod3ii
```


## Results
```{r}
(mod3results <- mod3 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height=80/25.4, fig.width=80/25.4}
mod3.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata3 <- mod3 %>% 
  emmeans(~SST, at = mod3.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod3_plot <- 
  ggplot(newdata3, aes(y = emmean, x = SST)) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
  geom_line() +
  geom_point(data=shark_daily_td,  aes(y=daily_min_temp, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
  theme_bjc() +
 # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  ylab("Daily min temp (°C)") + 
  xlab("\nSST (°C)"))
```

# Save
```{r fig.height=80/25.4, fig.width=80/25.4}
ggsave(mod3_plot, file = "output/mod3_plot.pdf", width = 100, height = 100, units = "mm", dpi = 300)
```




## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
(daily_min_temp_model_plot <- ggarrange(shark_daily_mintemp_plot, mod3_plot, ncol = 2, widths = c(2,1)))
ggsave(daily_min_temp_model_plot, file = "output/daily_min_temp_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)

```




##### Model 4: Daily Max Depth ~ SST #####

## Model formula
```{r}
mod4 <- glm(daily_max_depth ~ SST, data = shark_daily_td, family = gaussian)
mod4 <- gam(daily_max_depth ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod4 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod4 %>% performance::check_outliers()
mod4 %>% performance::r2()
```

## Results
```{r}
(mod4results <- mod4 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height=80/25.4, fig.width=80/25.4}
mod4.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata4 <- mod4 %>% 
  emmeans(~SST, at = mod4.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod4_plot <- 
  ggplot(newdata4, aes(y = emmean, x = SST)) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
  geom_line() +
  geom_point(data=shark_daily_td,  aes(y=daily_max_depth, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
  theme_bjc() +
 # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  ylab("Daily max depth (m)") + 
  xlab("\nSST (°C)"))
```


```{r fig.height=80/25.4, fig.width=80/25.4}
ggsave(mod4_plot, file = "output/mod4_plot.pdf", width = 100, height = 100, units = "mm", dpi = 300)
```



## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
(daily_max_depth_model_plot <- ggarrange(MaxDep_SST_plot, mod4_plot, ncol = 2, widths = c(2,1)))
ggsave(daily_max_depth_model_plot, file = "output/daily_max_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)

```

##### Model 5: Daily Mean Depth ~ SST #####


## Model formula
```{r}
mod5 <- glm(daily_mean_depth ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod5 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod5 %>% performance::check_outliers()
mod5 %>% performance::r2()
```

## Results
```{r}
(mod5results <- mod5 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height=80/25.4, fig.width=80/25.4}
mod5.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata5 <- mod5 %>% 
  emmeans(~SST, at = mod5.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod5_plot <- 
    ggplot(newdata5, aes(y = emmean, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_mean_depth, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    theme_bjc() +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Daily mean depth (m)") + 
    xlab("\nSST (°C)") +
    scale_y_continuous(limits = c(0,200)))

ggsave(mod5_plot, file = "output/mod5_plot.pdf", width = 100, height = 100, units = "mm", dpi = 300)
```



## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
(daily_mean_depth_model_plot <- ggarrange(shark_da, mod5_plot, ncol = 2, widths = c(2,1)))
ggsave(daily_mean_depth_model_plot, file = "output/daily_mean_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)

```



##### Model 6: Daily Min Depth ~ SST #####


## Model formula
```{r}
mod6 <- glm(daily_min_depth ~ SST, data = shark_daily_td, family = negative.binomial(2)) # nb capacity comes from MASS package
```

## Model validation
```{r}
# Dharma resids
mod6 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod6 %>% performance::check_outliers()
mod6 %>% performance::r2()
```

## Results
```{r}
(mod6results <- mod6 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r}
mod6.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata6 <- mod6 %>% 
  emmeans(~SST, at = mod6.grid) %>% 
  regrid() %>% 
  as_tibble()

(mod6_plot <- 
    ggplot(newdata6, aes(y = response, x = SST)) +
    geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), color=NA, alpha=0.3) +        # Had to change the ymin and max names as the model conducted in MASS
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_min_depth, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    theme_bjc() +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Daily min depth (m)") + 
    xlab("\nSST (°C)"))
ggsave(mod6_plot, file = "output/mod6_plot.pdf", width = 100, height = 100, units = "mm", dpi = 300)
```



## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
(daily_min_depth_model_plot <- ggarrange(minDep_SST_plot, mod6_plot, ncol = 2, widths = c(2,1)))
ggsave(daily_min_depth_model_plot, file = "output/daily_min_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)

```



```{r}
# END #
```






