---
title: "Shark daily depths ~ SST models"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	include = TRUE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
source("code/packages.R")
#rm(list=ls()) # Clear out environment if necessary
```

# Load data
```{r}
load(file = "data/Rdata/shark_daily_td.Rdata")
shark_daily_td
```


# We need models for:

4. Max daily depth ~ SST
5. Mean daily depth ~ SST
6. Min daily depth ~ SST

```{r}
## Check distributions so we know how to model
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_max_depth))
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_mean_depth))
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_min_depth))
shark_daily_td %>% ggplot() + geom_histogram(aes(x = daily_depth_sd))
```

All look approximately normal



##### Model 4: Daily Max Depth ~ SST #####

## Model formula
```{r}
mod4 <- glm(daily_max_depth ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod4 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod4 %>% performance::check_outliers()
mod4 %>% performance::r2()
```

## Results
```{r}
(mod4results <- mod4 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod4.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata4 <- mod4 %>% 
  emmeans(~SST, at = mod4.grid) %>% 
  regrid() %>% 
  as_tibble()

mod4_plot <- 
  ggplot(newdata4, aes(y = emmean, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_max_depth, x = SST), fill = "midnightblue", col = "black", alpha = 0.5, pch = 21) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), fill = "midnightblue", alpha=0.6) +
  geom_line(color = "black") +
  ylab("Daily max depth (m)") + 
  xlab("SST (째C)") +
  scale_y_continuous(limits = c(0, 210)) +
  theme_minimal() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) 

mod4_plot

ggsave(mod4_plot, file = "output/mod4_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```



## Add in relevant time-series data if required
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
#(daily_max_depth_model_plot <- ggarrange(MaxDep_SST_plot, mod4_plot, ncol = 2, widths = c(2,1)))
#ggsave(daily_max_depth_model_plot, file = "output/daily_max_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)
```

##### Model 5: Daily Mean Depth ~ SST #####


## Model formula
```{r}
mod5 <- glm(daily_mean_depth ~ SST, data = shark_daily_td, family = gaussian)
```

## Model validation
```{r}
# Dharma resids
mod5 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod5 %>% performance::check_outliers()
mod5 %>% performance::r2()
```

## Results
```{r}
(mod5results <- mod5 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod5.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata5 <- mod5 %>% 
  emmeans(~SST, at = mod5.grid) %>% 
  regrid() %>% 
  as_tibble()

mod5_plot <- 
  ggplot(newdata5, aes(y = emmean, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_mean_depth, x = SST), fill = "dodgerblue", col = "black", alpha = 0.6, pch = 21) +
  geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), fill = "dodgerblue", alpha=0.6) +
  geom_line(colour = "black") +
  ylab("Daily mean depth (m)") + 
  xlab("SST (째C)") +
  scale_y_continuous(limits = c(0,50)) +
  theme_minimal() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) 

mod5_plot

ggsave(mod5_plot, file = "output/mod5_plot.pdf", device = cairo_pdf, width = 56, height = 56, units = "mm", dpi = 300)
```



## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
#(daily_mean_depth_model_plot <- ggarrange(shark, mod5_plot, ncol = 2, widths = c(2,1)))
#ggsave(daily_mean_depth_model_plot, file = "output/daily_mean_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)
```



##### Model 6: Daily Min Depth ~ SST #####


## Model formula
```{r}
mod6 <- glm(daily_min_depth ~ SST, data = shark_daily_td, family = negative.binomial(2)) # nb capacity comes from MASS package
```

## Model validation
```{r}
# Dharma resids
mod6 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod6 %>% performance::check_outliers()
mod6 %>% performance::r2()
```

## Results
```{r}
(mod6results <- mod6 %>% tidy(conf.int=TRUE))
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod6.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata6 <- mod6 %>% 
  emmeans(~SST, at = mod6.grid) %>% 
  regrid() %>% 
  as_tibble()

mod6_plot <- 
  ggplot(newdata6, aes(y = response, x = SST)) +
  geom_point(data=shark_daily_td,  aes(y=daily_min_depth, x = SST), fill = "grey40", col = "black", alpha = 0.6, pch = 21) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), fill = "grey40", alpha=0.3) +        # Had to change the ymin and max names as the model conducted in MASS
  geom_line(colour = "black") +
  ylab("Daily min depth (m)") + 
  xlab("SST (째C)") +
  scale_y_continuous(limits = c(0,5)) +
  theme_minimal() +
  theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9)) 


mod6_plot

ggsave(mod6_plot, file = "output/mod6_plot.pdf", width = 56, height = 56, device = cairo_pdf, units = "mm", dpi = 300)
```



## Add in time-series data
```{r  fig.width = 180/25.4, fig.height = 60/25.4}
#(daily_min_depth_model_plot <- ggarrange(MinDep_SST_plot, mod6_plot, ncol = 2, widths = c(2,1)))
#ggsave(daily_min_depth_model_plot, file = "output/daily_min_depth_model_plot.pdf", width = 180, height = 60, units = "mm", dpi = 300)
```




##### Model 7: Mean daily depth SD ~ SST #####

## Model formula
Need a GAM


```{r}
mod7 <- gam(daily_depth_sd ~ s(SST, bs = "cs"), data = shark_daily_td, method = "REML")
mod7tw <- gam(daily_depth_sd ~ s(SST, bs = "cs"), data = shark_daily_td, family = tw, method = "REML")
AIC(mod7, mod7tw)
mod7 <- mod7tw
```

## Model validation
```{r}
# Dharma resids
mod7 %>% simulateResiduals(plot = TRUE) # Look ok apart from some quantile deviation
mod7 %>% gam.check()
mod7 %>% performance::check_outliers()
mod7 %>% performance::r2()
```

## Results
```{r}
(mod7results <- mod7 %>% tidy(conf.int=TRUE))
mod7 %>% summary()
```

## Plot results
```{r fig.height = 56/25.4, fig.width = 56/25.4}
mod7.grid <- shark_daily_td %>% 
  with(list(SST = seq_range(SST, n = 100)))

newdata7 <- mod7 %>% 
  emmeans(~SST, at = mod7.grid) %>% 
  regrid() %>% 
  as_tibble()

mod7_plot <- 
    ggplot(newdata7, aes(y = emmean, x = SST)) +
    geom_ribbon(aes(ymin=lower.CL, ymax=upper.CL), color=NA, alpha=0.3) +
    geom_line() +
    geom_point(data=shark_daily_td,  aes(y=daily_depth_sd, x = SST), fill = "grey40", col = "black", alpha = 0.4, pch = 21) +
    # theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
    ylab("Std. dev. daily depth (m)") + 
    xlab("SST (째C)") +
    theme_minimal() +
    theme(#plot.background = element_rect(colour = "black", linewidth = 1),
        axis.title.y = element_text(size = 9),
        axis.title.x = element_text(size = 9))


mod7_plot

ggsave(mod7_plot, file = "output/mod7_plot.pdf", width = 62, height = 62, device = cairo_pdf, units = "mm", dpi = 300)
?mgcv
citation("mgcv")
```


```{r}
# ENDS
```




