---
title: "Results output for GLMM1"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
#rm(list=ls()) # Clear out environment if necessary
source("code/packages.R")
```


# Introducion # 

This document digs down into the interaction terms of the first mixed effect model:

mod1 = daily_movt ~ SST * Season 

We are interested in the following interactions:

SST * Season - from Aim 3 in the intro

Also a reminder of ranges/values of some variables:

# SST (annual)
min       median    mean    max
25.67     28.56     28.34   30.53

# SST (seasonal)
min.  max.  Season
25.67	27.81	Winter		
28.08	30.53	Summer		
26.49	29.66	Spring		
27.74	30.09	Autumn

We'll use these values to set sensible limits and contrasts on our model data.

# Firt load model and data
```{r message=FALSE, warning=FALSE, include=FALSE}
sourceRmd("code/02_model/glmms_fit.Rmd")
```

# Season * SST

# Create reference grid
```{r}
mod1_grid <- shark_individual_daily_td %>% 
    with(list(Season = levels(Season),
              SST = seq_range(SST, n = 100))) # So we'll get 100 emmeans estimates per season
```

# Use emmmeans to create newdata tbl
```{r}
# New data  
newdata_mod1 <- mod1 %>% 
  emmeans(~Season|SST, at = season_SST_grid) %>% 
  regrid() %>% 
  as_tibble()
```

# We need to curtail the model output so that it only provides estimates within or close to the range of the actual observed data
But also want to plot the whole model in the background, so need to get the full range out of the newdata:
```{r}
# Get the full range to plot 
full_response <- newdata_mod1
```

# Reminder of actual seasonal ranges
min   max   Season
25.67	27.81	Winter		
28.08	30.53	Summer		
26.49	29.66	Spring		
27.74	30.09	Autumn

# Truncate
```{r}
newdata_mod1 <- newdata_mod1 %>% 
  filter(Season == "Spring" & SST >= 26.49 & 
         Season == "Spring" & SST <= 29.66 |
         Season == "Autumn" & SST >= 27.74 &
         Season == "Autumn" & SST <= 30.09 |
         Season == "Summer" & SST >= 28.08 |
         Season == "Winter" & SST <= 27.81)
```

Just need to make some labels for 
```{r}
Seasons <- as_tibble(list(Row = seq(1,4),
                          Season = c("Spring", "Summer", "Autumn", "Winter"))) %>% 
  mutate(Season = fct_relevel(Season, c("Spring", "Summer", "Autumn", "Winter")))
  

```


# And plot
```{r  fig.width = 65/25.4, fig.height = 105/25.4}
seasonSST_plot <- 
newdata_mod1 %>% 
  ggplot() +
  geom_point(data = shark_individual_daily_td, aes(x = SST, y = daily_movement, colour = Season, fill = Season), 
             pch = 20, alpha = 0.05, position = position_jitter(width = 0.5, height = 0.5)) +
  geom_ribbon(data = full_response, aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, fill = Season, group = Season), alpha = 0.2) +
  geom_line(data = full_response, aes(x = SST, y = response, colour = Season, group = Season), alpha = 0.5, size = 0.5) +
  geom_ribbon(aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, fill = Season, group = Season), alpha = 0.5) +
  geom_line(aes(x = SST, y = response, colour = Season, group = Season), size = 1) +
  #scale_shape_manual(values = c(21, 22, 23)) +
  scale_fill_manual(values = c("seagreen","red", "orange", "darkblue")) +
  scale_colour_manual(values = c("seagreen","red", "orange", "darkblue")) +
  theme_minimal() +
  scale_y_continuous(limits = c(10, 115), breaks = c(0, 25, 50, 75, 100),  expand = c(0,0)) +
  scale_x_continuous(limits = c(25.65, 30.55), expand = c(0,0)) +
  theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 0.5)) +
  ylab("Daily depth movement (m ± 95%CI)") + 
  xlab("SST (°C)") +
  theme(legend.position = "none",
        strip.text = element_blank(),
        #legend.position = c(0.16, 0.7),
        #legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 8, colour = "black")) +
        #legend.title = element_text(size = 8),
        #legend.text = element_text(size = 8)) +
  facet_wrap(~Season, ncol = 1) +
  geom_text(data = Seasons, aes(x=26, y=100, label = Season), hjust = 0.2, vjust = -0.2, size = 3)
seasonSST_plot
```

# And plot (Gaussian model)
```{r  fig.width = 65/25.4, fig.height = 105/25.4}
  newdata_seasonSST %>% 
  ggplot() +
  geom_point(data = shark_individual_daily_td, aes(x = SST, y = daily_movement, colour = Season, fill = Season), 
             pch = 20, alpha = 0.05, position = position_jitter(width = 0.5, height = 0.5)) +
  geom_ribbon(data = full_response, aes(x = SST, y = emmean, ymin=lower.CL, ymax=upper.CL, fill = Season, group = Season), alpha = 0.2) +
  geom_line(data = full_response, aes(x = SST, y = emmean, colour = Season, group = Season), alpha = 0.5, size = 0.5) +
  geom_ribbon(aes(x = SST, y = emmean, ymin=lower.CL, ymax=upper.CL, fill = Season, group = Season), alpha = 0.5) +
  geom_line(aes(x = SST, y = emmean, colour = Season, group = Season), size = 1) +
  #scale_shape_manual(values = c(21, 22, 23)) +
  scale_fill_manual(values = c("seagreen","red", "orange", "darkblue")) +
  scale_colour_manual(values = c("seagreen","red", "orange", "darkblue")) +
  theme_minimal() +
  scale_y_continuous(limits = c(10, 115), breaks = c(0, 25, 50, 75, 100),  expand = c(0,0)) +
  scale_x_continuous(limits = c(25.65, 30.55), expand = c(0,0)) +
  theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 0.5)) +
  ylab("Daily depth movement (m ± 95%CI)") + 
  xlab("SST (°C)") +
  theme(legend.position = "none",
        strip.text = element_blank(),
        #legend.position = c(0.16, 0.7),
        #legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 8, colour = "black")) +
  #legend.title = element_text(size = 8),
  #legend.text = element_text(size = 8)) +
  facet_wrap(~Season, ncol = 1) +
  geom_text(data = Seasons, aes(x=26, y=100, label = Season), hjust = 0.2, vjust = -0.2, size = 3)
```


# Save
```{r}
ggsave(seasonSST_plot, file = "output/seasonSST_plot.pdf", width = 65, height = 105, device = cairo_pdf, units = "mm", dpi = 300)
```

# Actual results

## Reminder of R2 and summary
```{r}
mod1 %>% performance::r2_nakagawa()
mod1 %>% summary()

```

# emmeans
```{r}
mod1 %>%  
  emmeans(specs = pairwise ~ Season, infer = TRUE, type = "response") %>% 
  confint()
```

# Trend

```{r}
mod1 %>% tidy(conf.int = TRUE)
```

```{r}
#season_trends <- 
  mod1 %>% 
  emtrends(specs = pairwise~"Season", var = "SST", infer = TRUE) %>% 
  regrid()

season_trendsum <- summary(season_trends, infer = TRUE)

season_trendsum

mod1 %>% confint() 
```

```{r}
library(modelbased)
mod1 %>% estimate_slopes(trend = "SST", at = "Season")
mod1 %>% ggeffect()
```


```{r}
# END #
```


