---
title: "Individual movement results - Season * SST"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
#rm(list=ls()) # Clear out environment if necessary
source("code/packages.R")
```


# Load data
```{r}
load(file = "data/Rdata/shark_individual_daily_td.Rdata")
```


# Load model
```{r}
load(file = "data/Rdata/final_model.Rdata")
```

# Intro
This document digs down into the interactive terms in the model:

Season * SST
Sex * TL

## To do this we'll need to calculate mean levels of:
- SST
- TL
and perform contrast across those levels

# SST = approx 28.5°C
```{r}
shark_individual_daily_td %>% 
  ungroup() %>% 
  summarise_at(vars(SST),list(min = min, median=median, mean = mean, max = max))
```
# TL - approx 150cm
```{r}
shark_individual_daily_td %>% 
  ungroup() %>% 
  group_by(Sex) %>% 
  summarise_at(vars(TL),
               list(min=min, 
                    mean = mean,
                    max=max))
```

# Calculate trends at those levels across:
1°C 
10 cm TL

## Create reference grid, emmmeans tbl and truncate SSTs as above
```{r}
# Reference grid
results_grid <- shark_individual_daily_td %>% 
    with(list(Season = levels(Season),
              TL = c(155, 145),                       # Mean TL = 150cm
              Sex = levels(Sex),
              SST = c(29, 28)))                       # Mean SST = 28.34, Median SST = 28.56
```



```{r}
season_sst_slopes <- 
  sst_maxd_final %>% 
  emmeans(~SST|Season, at = results_grid) %>%
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
 # tab_df()
  as_tibble() %>% 
  mutate(across(where(is.numeric), round, 2))

season_sst_slopes
```


```{r}
sex_tl_slopes <- 
  sst_maxd_final %>% 
  emmeans(~TL|Sex, at = results_grid) %>%
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
 # tab_df()
  as_tibble() %>% 
  mutate(across(where(is.numeric), round, 2))
  
sex_tl_slopes
```




# And plot
```{r  fig.width = 100/25.4, fig.height = 80/25.4}
(Season_SST_plot <- 
newdata_Season_SST %>% 
  ggplot() +
  geom_linerange(aes(x = Season, y = response, ymin=asymp.LCL, ymax=asymp.UCL, colour = SST, group = SST),  position=position_dodge(0.25)) +
  geom_point(aes(x = Season, y = response, colour = SST, fill = SST), position=position_dodge(0.25), size = 4) +
  #scale_shape_manual(values = c(21, 22, 23)) +
  scale_fill_manual(values = c("darkblue", "dodgerblue", "red")) +
  scale_colour_manual(values = c("darkblue", "dodgerblue", "red")) +
  theme_minimal() +
  scale_y_continuous(limits = c(10, 115), breaks = c(0, 25, 50, 75, 100)) +
  #theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  #ylab("Daily depth movement (m)\n(e.m. mean ± 95%CI)\n") + 
  xlab("\nSeason") +
  theme(legend.position = c(0.88, 0.74 ),
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.title = element_text(size = 8)))
        
```

# Save
```{r}
ggsave(Season_SST_plot, file = "output/Season_SST_plot.pdf", width = 80, height = 90, units = "mm", dpi = 300)
```


# 1.a. Calculate mean effect of 1°C SST change on depth use across seasons (i.e. slope for all seasons)
## Create reference grid, emmmeans tbl and truncate SSTs as above
```{r}
# Reference grid
Season_SST_grid_1deg <- shark_individual_daily_td %>% 
    with(list(Season = levels(Season),
              TL = c(145, 155),                       # Mean TL = 150cm
#              Sex = levels(Sex),
             SST = c(28, 29)))                       # Mean SST = 28.34, Median SST = 28.56
   
# New data
newdata_Season_SST_1deg <- sst_maxd_final %>% 
  emmeans(~Season|SST, at = Season_SST_grid_1deg) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(SST = factor(SST))
```



```{r}
#season_emm <- 
  sst_maxd_final %>% 
  emmeans(~SST|Season, at = Season_SST_grid_1deg) %>%
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  tab_df()
  as_tibble() 
```







2. SST as continuous

# Create reference grid, emmmeans tbl and truncate SSTs as above
```{r}
# Reference grid
Season_SST_grid2 <- shark_individual_daily_td %>% 
    with(list(Season = levels(Season),
              SST = seq_range(SST, n = 100)))

# New data  
newdata_Season_SST2 <- sst_maxd_final %>% 
  emmeans(~Season|SST, at = Season_SST_grid2) %>% 
  regrid() %>% 
  as_tibble()

# Reminder of actual seasonal ranges
#min    max   Season
# 25.67	27.81	Winter		
# 28.08	30.53	Summer		
# 26.49	29.66	Spring		
# 27.74	30.09	Autumn

# Truncate
newdata_Season_SST2 <- newdata_Season_SST2 %>% 
  filter(Season == "Spring" & SST >= 26.49 & 
         Season == "Spring" & SST <= 29.66 |
         Season == "Autumn" & SST >= 27.74 &
         Season == "Autumn" & SST <= 30.09 |
         Season == "Summer" & SST >= 28.08 |
         Season == "Winter" & SST <= 27.81)
```

# And plot
```{r  fig.width = 150/25.4, fig.height = 70/25.4}
(Season_SST_plot2 <- 
newdata_Season_SST2 %>% 
  ggplot() +
  geom_ribbon(aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, group = Season), alpha = 0.2) +
  geom_line(aes(x = SST, y = response, colour = Season, group = Season), size = 1) +
  #scale_shape_manual(values = c(21, 22, 23)) +
  scale_fill_manual(values = c("dodgerblue","red", "orange", "darkblue")) +
  scale_colour_manual(values = c("dodgerblue","red", "orange", "darkblue")) +
  theme_minimal() +
  scale_y_continuous(limits = c(10, 115), breaks = c(0, 25, 50, 75, 100)) +
  theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 0.5)) +
  ylab("Daily depth movement (m)\n(e.m. mean ± 95%CI)") + 
  xlab("\nSST (°C)") +
  theme(legend.position = c(0.16, 0.7),
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)))

```

# Save
```{r}
ggsave(Season_SST_plot2, file = "output/Season_SST_plot2.pdf", width = 140, height = 70, device = cairo_pdf, units = "mm", dpi = 300)
```

```{r  fig.width = 180/25.4, fig.height = 90/25.4}
Season_SST_fig <- cowplot::plot_grid(Season_SST_plot2, Season_SST_plot, nrow = 1, labels = c('A', 'B'), label_size = 9, align = "h", rel_widths = c(1.1, 0.6)) +
  theme(plot.background = element_rect(color = "black"))
Season_SST_fig
```

# Save
```{r}
ggsave(Season_SST_plot3, file = "output/Season_SST_plot3.pdf", width = 170, height = 70, units = "mm", dpi = 300)
ggsave(Season_SST_fig, file = "output/Season_SST_fig.pdf", width = 170, height = 70, device = cairo_pdf, units = "mm", dpi = 300)
```


# Extract trend

```{r}
sst_maxd_final %>% 
  tidy(conf.int = TRUE)
```

```{r}
#season_trends <- 
  sst_maxd_final %>% 
  emtrends(specs = pairwise~"Season", var = "SST", infer = TRUE)

season_trendsum <- summary(season_trends, infer = TRUE)

season_trendsum
```

```{r}
sst_maxd_final %>% 
  estimate_slopes(trend = "SST", at = "Season")

sst_maxd_final %>% 
  ggeffect()
```


```{r}
# END #
```