---
title: "Individual movement results - SST * Sex"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
rm(list=ls()) # Clear out environment if necessary
source("code/packages.R")
```


This document digs down into the interactive effect of SST and Sex on the magnitude of daily depth movements in individual sharks

# Load
```{r}
load(file = "data/Rdata/shark_individual_daily_td.Rdata")
```

# Load model
```{r}
load(file = "data/Rdata/final_model.Rdata")
```


# Options

Probably want two plots here -

1. Differences in depth use between sexes at the average annual SST
2. Differences in depth use between sexes across the full range of SSTs

1. Inter-sex differences at average SST:

```{r}
shark_individual_daily_td %>% 
  ungroup %>% 
  summarise(mean_SST = mean(SST))
  
```

# Create reference grid, emmmeans tbl and truncate SSTs as above
```{r}
# Reference grid
sex_SST_grid <- shark_individual_daily_td %>% 
    with(list(Sex = levels(Sex),
              SST = 28.34))
# New data
newdata_sex_SST <- sst_maxd_final %>% 
  emmeans(~Sex|SST, at = sex_SST_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(SST = factor(SST))
```


# And plot
```{r  fig.width = 100/25.4, fig.height = 80/25.4}
(sex_SST_plot1 <- 
newdata_sex_SST %>% 
  ggplot() +
  geom_linerange(aes(x = Sex, y = response, ymin=asymp.LCL, ymax=asymp.UCL), colour = "grey40") +
  geom_point(aes(x = Sex, y = response, shape = Sex, fill = Sex), size = 4) +
  scale_shape_manual(values = c(23, 22))  +
  scale_fill_manual(values = c("darkblue", "dodgerblue")) +
  scale_colour_manual(values = c("darkblue", "dodgerblue")) +
  theme_minimal() +
# theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  ylab("Daily depth movement (m)\n(e.m. mean ± 95%CI)\n") + 
  xlab("\nSex") +
  theme(legend.position = c(0.25, 0.82),
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 9),
        legend.title = element_text(size = 8)))
```

# Save
```{r}
ggsave(sex_SST_plot1, file = "output/Sex_SST_plot1.pdf", width = 80, height = 90, units = "mm", dpi = 300)
```



2. Inter-sex differences across full range of SST:

# Create reference grid, emmmeans tbl and truncate SSTs as above
```{r}
# Reference grid
Sex_SST_grid2 <- shark_individual_daily_td %>% 
    with(list(Sex = levels(Sex),
              SST = seq_range(SST, n = 100)))

# New data  
newdata_Sex_SST2 <- sst_maxd_final %>% 
  emmeans(~Sex|SST, at = Sex_SST_grid2) %>% 
  regrid() %>% 
  as_tibble()
```

# And plot
```{r  fig.width = 140/25.4, fig.height = 60/25.4}
(Sex_SST_plot2 <- 
newdata_Sex_SST2 %>% 
  ggplot() +
  geom_ribbon(aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, group = Sex), alpha = 0.2) +
  geom_line(aes(x = SST, y = response, colour = Sex, group = Sex), size = 1) +
  #scale_shape_manual(values = c(21, 22, 23)) +
  scale_fill_manual(values = c("darkblue", "dodgerblue")) +
  scale_colour_manual(values = c("darkblue", "dodgerblue")) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 110), breaks = c(0, 25, 50, 75, 100)) +
# theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  ylab("Daily depth movement (m)\n(e.m. mean ± 95%CI)") + 
  xlab("\nSST (°C)") +
  theme(legend.position = c(0.14, 0.76),
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8)))
```

# Save
```{r}
ggsave(Sex_SST_plot2, file = "output/Sex_SST_plot2.pdf", width = 115, height = 60, units = "mm", dpi = 300)
```


```{r}
# END #
```