---
title: "Individual movement results - Season * SST"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
#rm(list=ls()) # Clear out environment if necessary
source("code/packages.R")
```


This document digs down into the interactive effect of Season and SST on daily depth movments in individual sharks
The results of the model will make predictions for the whole range of SSTs for all four seasons so we need to truncate our output model data accordingly.
To see the real range in SSTs across the seasons we'll need to load the base tbl and check:

# Load data
```{r}
load(file = "data/Rdata/shark_individual_daily_td.Rdata")
```


# Load model
```{r}
load(file = "data/Rdata/final_model.Rdata")
```


# Options
Can present these results two ways: 
1. With SST as an ordinal variable using max, min and median values within the observed range - arguably easier to interpret
2. With SST as a continuous variable presenting all the predicted values with a confint 'ribbon' - potentially 'truer' to the data

Will do both and present together

1. SST as ordinal

Determine IQR
```{r}
shark_individual_daily_td %>% 
  ungroup() %>% 
  summarise_at(vars(SST),
               list(min=min, 
                    Q1=~quantile(., probs = 0.25),
                    median=median, 
                    mean = mean,
                    Q3=~quantile(., probs = 0.75),
                    max=max))
```
# First need to relevel season
```{r}


```



# Also need to know the seasonal ranges
```{r}
season_ranges <-  
  shark_individual_daily_td %>% 
  filter(Season == "Winter") %>%  # 25.7 - 27.8°C
  droplevels() %>% 
  summarise_at(vars(SST),
               list(min=min, max=max)) %>%
  mutate(Season = "Winter") %>% 
  bind_rows(
  shark_individual_daily_td %>% 
  filter(Season == "Summer") %>%  # 28.1 - 30.5°C
  droplevels() %>% 
  summarise_at(vars(SST),
               list(min=min, max=max)) %>% 
  mutate(Season = "Summer")) %>% 
  bind_rows(
  shark_individual_daily_td %>% 
  filter(Season == "Spring") %>%  # 26.5 - 29.7°C
  droplevels() %>% 
  summarise_at(vars(SST),
               list(min=min, max=max)) %>% 
  mutate(Season = "Spring")) %>% 
  bind_rows(
  shark_individual_daily_td %>% 
  filter(Season == "Autumn") %>%  # 27.7 - 30.1°C
  droplevels() %>% 
  summarise_at(vars(SST),
               list(min=min, max=max)) %>% 
  mutate(Season = "Autumn"))
```

min.  max.  Season
25.67	27.81	Winter		
28.08	30.53	Summer		
26.49	29.66	Spring		
27.74	30.09	Autumn


We'll use these values to set sensible limits on our model data


# Create reference grid, emmmeans tbl and truncate SSTs as above
```{r}
# Reference grid
Season_SST_grid <- shark_individual_daily_td %>% 
    with(list(Season = levels(Season),
             # SST = c(25.7, 28.6, 30.5)))
             SST = c(26, 28, 30)))

# New data
newdata_Season_SST <- sst_maxd_final %>% 
  emmeans(~Season|SST, at = Season_SST_grid) %>% 
  regrid() %>% 
  as_tibble()

# Truncate SST values according to season
newdata_Season_SST <- newdata_Season_SST %>% 
  filter(Season == "Spring" | #& SST > 27.5 | 
         Season == "Autumn" | #& SST > 26.5 |
         Season == "Summer" & SST > 28.8 |
         Season == "Winter" & SST < 27.9) %>% 
  mutate(SST = factor(SST))
```

# Contrasts
```{r}
#season_emm <- 
  sst_maxd_final %>% 
  emmeans(~SST|Season, at = Season_SST_grid) %>%
  regrid() %>% 
  pairs() %>% 
  confint() %>% 
  as_tibble() %>% 
  filter(!Season == "Summer" | contrast == "SST28 - SST30",
         !Season == "Winter" | contrast == "SST26 - SST28")
```




# And plot
```{r  fig.width = 100/25.4, fig.height = 80/25.4}
(Season_SST_plot <- 
newdata_Season_SST %>% 
  ggplot() +
  geom_linerange(aes(x = Season, y = response, ymin=asymp.LCL, ymax=asymp.UCL, colour = SST, group = SST),  position=position_dodge(0.25)) +
  geom_point(aes(x = Season, y = response, colour = SST, fill = SST), position=position_dodge(0.25), size = 4) +
  #scale_shape_manual(values = c(21, 22, 23)) +
  scale_fill_manual(values = c("darkblue", "dodgerblue", "red")) +
  scale_colour_manual(values = c("darkblue", "dodgerblue", "red")) +
  theme_minimal() +
  scale_y_continuous(limits = c(10, 115), breaks = c(0, 25, 50, 75, 100)) +
  #theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  #ylab("Daily depth movement (m)\n(e.m. mean ± 95%CI)\n") + 
  xlab("\nSeason") +
  theme(legend.position = c(0.88, 0.74 ),
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.title = element_text(size = 8)))
        
```

# Save
```{r}
ggsave(Season_SST_plot, file = "output/Season_SST_plot.pdf", width = 80, height = 90, units = "mm", dpi = 300)
```


2. SST as continuous

# Create reference grid, emmmeans tbl and truncate SSTs as above
```{r}
# Reference grid
Season_SST_grid2 <- shark_individual_daily_td %>% 
    with(list(Season = levels(Season),
              SST = seq_range(SST, n = 100)))

# New data  
newdata_Season_SST2 <- sst_maxd_final %>% 
  emmeans(~Season|SST, at = Season_SST_grid2) %>% 
  regrid() %>% 
  as_tibble()

# Get the full range to plot 
full_response <- newdata_Season_SST2

# Reminder of actual seasonal ranges
#min    max   Season
# 25.67	27.81	Winter		
# 28.08	30.53	Summer		
# 26.49	29.66	Spring		
# 27.74	30.09	Autumn

# Truncate
newdata_Season_SST2 <- newdata_Season_SST2 %>% 
  filter(Season == "Spring" & SST >= 26.49 & 
         Season == "Spring" & SST <= 29.66 |
         Season == "Autumn" & SST >= 27.74 &
         Season == "Autumn" & SST <= 30.09 |
         Season == "Summer" & SST >= 28.08 |
         Season == "Winter" & SST <= 27.81)
```

# And plot
```{r  fig.width = 72/25.4, fig.height = 150/25.4}
Season_SST_plot2 <- 
newdata_Season_SST2 %>% 
  ggplot() +
  geom_point(data = shark_individual_daily_td, aes(x = SST, y = daily_movement, colour = Season, fill = Season), 
             pch = 20, alpha = 0.05, position = position_jitter(width = 0.5, height = 0.5)) +
  geom_ribbon(data = full_response, aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, fill = Season, group = Season), alpha = 0.2) +
  geom_line(data = full_response, aes(x = SST, y = response, colour = Season, group = Season), alpha = 0.5, size = 0.5) +
  geom_ribbon(aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, fill = Season, group = Season), alpha = 0.5) +
  geom_line(aes(x = SST, y = response, colour = Season, group = Season), size = 1) +
  #scale_shape_manual(values = c(21, 22, 23)) +
  scale_fill_manual(values = c("seagreen","red", "orange", "darkblue")) +
  scale_colour_manual(values = c("seagreen","red", "orange", "darkblue")) +
  theme_minimal() +
  scale_y_continuous(limits = c(10, 115), breaks = c(0, 25, 50, 75, 100),  expand = c(0,0)) +
  scale_x_continuous(limits = c(25.65, 30.55), expand = c(0,0)) +
  theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 0.5)) +
  ylab("Daily depth movement (m ± 95%CI)") + 
  xlab("SST (°C)") +
  theme(legend.position = "none",
        strip.text = element_text(angle = 90, vjust = -0.5, size = 8),
        strip.placement = "outside",
        #legend.position = c(0.16, 0.7),
        #legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 8),
        axis.text = element_text(size = 8)) +
        #legend.title = element_text(size = 8),
        #legend.text = element_text(size = 8)) +
  facet_wrap(~Season, ncol = 1, strip.position = "right")
Season_SST_plot2
```

# Save
```{r}
ggsave(Season_SST_plot2, file = "output/Season_SST_plot2.pdf", width = 80, height = 160, device = cairo_pdf, units = "mm", dpi = 300)
```

```{r  fig.width = 180/25.4, fig.height = 90/25.4}
Season_SST_fig <- cowplot::plot_grid(Season_SST_plot2, Season_SST_plot, nrow = 1, labels = c('A', 'B'), label_size = 9, align = "h", rel_widths = c(1.1, 0.6)) +
  theme(plot.background = element_rect(color = "black"))
Season_SST_fig
```

# Save
```{r}
ggsave(Season_SST_plot3, file = "output/Season_SST_plot3.pdf", width = 170, height = 70, units = "mm", dpi = 300)
ggsave(Season_SST_fig, file = "output/Season_SST_fig.pdf", width = 170, height = 70, device = cairo_pdf, units = "mm", dpi = 300)
```


# Extract trend

```{r}
sst_maxd_final %>% 
  tidy(conf.int = TRUE)
```

```{r}
#season_trends <- 
  sst_maxd_final %>% 
  emtrends(specs = pairwise~"Season", var = "SST", infer = TRUE)

season_trendsum <- summary(season_trends, infer = TRUE)

season_trendsum
```

```{r}
sst_maxd_final %>% 
  estimate_slopes(trend = "SST", at = "Season")

sst_maxd_final %>% 
  ggeffect()
```


```{r}
# END #
```