---
title: "Results output for GLMM2"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
#rm(list=ls()) # Clear out environment if necessary
source("code/packages.R")
```


# Introducion # 

This document digs down into the interaction terms of the second mixed effect models we have.

mod2 = daily_movt ~ SST * Sex * TL (from Aim 4 in the intro)

From this model we are interested in the following interactions:

SST * TL - from Aim 4 in the intro
SST * Sex - from Aim 4 in the intro
TL * Sex - not explicitly a stated aim but we have the interaction so we'll investigate anyways.

# TL - we have a bimodal distribution with a large gap between 141 and 161 cm for both sexes. Otherwise the size range is similar
Female  125-176 cm
Male    130-171 cm

We'll use these values to set sensible limits and contrasts on our model data.

# Firt load model and data
```{r message=FALSE, warning=FALSE, include=FALSE}
sourceRmd("code/02_model/glmms_fit.Rmd")
```

# Interaction: TL * SST 

# Create reference grid
```{r}
mod2_grid <- shark_individual_daily_td %>% 
    with(list(TL = c(125, 175),
              Sex = levels(Sex),
              SST = seq_range(SST, n = 100))) # So we'll get 100 emmeans estimates per season
```

# Use emmmeans to create newdata tbl
```{r}
# New data  
newdata_mod2 <- mod2 %>% 
  emmeans(~TL|SST|Sex, at = mod2_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(T.L. = as.factor(case_when(TL < 145 ~ "Small",
                                      TL > 150 ~ "Large")))
```

# Make plotting variable in data
```{r}
shark_individual_daily_td <- 
  shark_individual_daily_td %>% 
  mutate(T.L. = as.factor(case_when(TL < 145 ~ "Small",
                                      TL > 150 ~ "Large")))
```



# And plot
```{r  fig.width = 100/25.4, fig.height = 100/25.4}
sexTL_SST_fig <- 
newdata_mod2 %>% 
  ggplot() +
  geom_point(data = shark_individual_daily_td, aes(x = SST, y = daily_movement, fill = T.L., colour = T.L.), 
            pch = 19, alpha = 0.1, position = position_jitter(width = 0.5, height = 0.5)) +
  geom_ribbon(aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, fill = T.L., group = T.L.), alpha = 0.5) +
  geom_line(aes(x = SST, y = response, group = T.L.), colour = "brown4", size = 0.3) +
  scale_fill_manual(values = c("#3A6B35", "#E3B448")) +
  scale_colour_manual(values = c("#3A6B35", "#E3B448")) +
  #scale_color_viridis_d(option = "-D") +
  #scale_fill_viridis_d(option = "-D") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 160), breaks = c(0, 25, 50, 75, 100, 125, 150),  expand = c(0,0)) +
  scale_x_continuous(limits = c(25.65, 30.55), expand = c(0,0)) +
  #theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 0.5)) +
  ylab("Daily depth range (m)") + 
  xlab("SST (°C)") +
  theme(legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(size = 8, colour = "black"),
        strip.text = element_text(size = 9, colour = "black"),
        legend.position = c(0.02, 0.98),
        legend.justification = c(0,1),
        legend.key.size = unit(0.35, "cm"),
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 8, colour = "black")) +
        #legend.title = element_text(size = 8),
        #legend.text = element_text(size = 8)) +
  #geom_text(data = Seasons, aes(x=26, y=100, label = Season), hjust = 0.2, vjust = -0.2, size = 3) +
  facet_wrap(~Sex, ncol = 1, strip.position = "right")
TL_sex_SST_fig
```


# Plot with facets other way around 
```{r  fig.width = 100/25.4, fig.height = 60/25.4}
TL_sex_SST_fig <- 
newdata_mod2 %>% 
  ggplot() +
  geom_point(data = shark_individual_daily_td, aes(x = SST, y = daily_movement, fill = Sex, colour = Sex), 
            pch = 21, alpha = 0.05, position = position_jitter(width = 0.5, height = 0.5)) +
  geom_ribbon(aes(x = SST, y = response, ymin=asymp.LCL, ymax=asymp.UCL, fill = Sex, group = Sex), alpha = 0.5) +
  geom_line(aes(x = SST, y = response, group = Sex), colour = "black", size = 0.3) +
  scale_fill_manual(values = c("#440154", "#5ec962")) +
  scale_colour_manual(values = c("#440154", "#5ec962")) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 160), breaks = c(0, 50, 100, 150),  expand = c(0.1,0.1)) +
  scale_x_continuous(limits = c(25.6, 30.6), breaks = c(26,28,30), expand = c(0,0)) +
  #theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 0.5)) +
  ylab("Daily depth range (m)") + 
  xlab("SST (°C)") +
  theme(#legend.direction = "horizontal",
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, colour = "black"),
        strip.text = element_blank(),
        legend.position = "none",
        #legend.justification = c(0,1),
        #legend.key.size = unit(0.35, "cm"),
        #legend.background = element_rect(fill = "white", colour = "black"),
        axis.title = element_text(size = 9, colour = "black"),
        axis.text = element_text(size = 8, colour = "black")) +
        #legend.title = element_text(size = 8),
        #legend.text = element_text(size = 8)) +
  #geom_text(data = Seasons, aes(x=26, y=100, label = Season), hjust = 0.2, vjust = -0.2, size = 3) +
  facet_grid(vars(Sex), vars(T.L.))

TL_sex_SST_fig
```


```{r}
emm_demog <- 
  mod2 %>% 
  emmeans(~TL|Sex, at = mod2_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(T.L. = as.factor(case_when(TL < 145 ~ "Small",
                                      TL > 150 ~ "Large")))

```

```{r}
shark_individual_daily_td  %$% 
  summary(daily_movement)
```


```{r  fig.width = 100/25.4, fig.height = 60/25.4}
TL_sex_fig <- 
  newdata_mod2 %>% 
  ggplot() +
  geom_point(data = shark_individual_daily_td, aes(x = T.L., y = daily_movement, colour = Sex, group = Sex), alpha = 0.01, pch = 15, 
             size = 3, position = position_jitter(seed = 1, width = 0.35, height = 0.1)) +
  geom_point(aes(x = T.L., y = response,  colour = Sex, group = Sex), fill = "transparent", alpha = 0.5, pch = 21, 
             position = position_jitter(seed = 1, width = 0.35, height = 0.1)) +
  geom_point(aes(x = T.L., y = response, fill = Sex, group = Sex), colour = "transparent", alpha = 0.3, pch = 21, 
             position = position_jitter(seed = 1, width = 0.35, height = 0.1)) +
  geom_pointrange(data = emm_demog, aes(x = T.L., y = response, ymin=asymp.LCL, ymax=asymp.UCL, shape = Sex, group = Sex), colour = "black", fill = "transparent", #pch = 22,
                  alpha = 1,  size = 0.8, position = position_dodge(width = 0.35)) +
  geom_pointrange(data = emm_demog, aes(x = T.L., y = response, ymin=asymp.LCL, ymax=asymp.UCL, shape = Sex, fill = Sex, group = Sex), colour = NA, # pch = 22,
                  alpha = 0.8,  size = 0.8, position = position_dodge(width = 0.35)) +
  scale_fill_manual(values = c("#440154", "#5ec962")) +       # viridis cols
  scale_colour_manual(values = c("#440154", "#5ec962")) +      # viridis cols
  scale_shape_manual(values = c(25, 24)) +
  #scale_color_viridis(discrete = TRUE) +
  #scale_fill_viridis(discrete = TRUE) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 150), breaks = c(0, 50, 100, 150),  expand = c(0,0)) +
  scale_x_discrete(expand = c(0.1, 0.1)) +
  #theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 0.5)) +
  ylab("Daily dep rg (m)") + 
  xlab("Body size") +
  theme(strip.text = element_text(size = 9, colour = "black"),
        #legend.direction = "horizontal",
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, colour = "black"),
        #legend.position = c(0.5, 0.98),
        legend.position = "none",
        #legend.justification = c(0.5,1),
        #legend.key.size = unit(0.35, "cm"),
        #legend.background = element_rect(fill = "white", colour = "black"),
        axis.title.x = element_text(size = 9, colour = "black"),
        axis.title.y = element_text(size = 9, hjust = 0, colour = "black"),
        axis.text = element_text(size = 8, colour = "black"))

TL_sex_fig 
```

```{r}
male <- readPNG("data/male.png") %>% grid::rasterGrob(interpolate = TRUE)
female <- readPNG("data/female.png") %>% grid::rasterGrob(interpolate = TRUE)
```




# Plot together

```{r  fig.width = 90/25.4, fig.height = 110/25.4}
TL_sex_SST_figcombined <- 
  cowplot::plot_grid(TL_sex_SST_fig, TL_sex_fig, nrow = 2, rel_heights = c(1,0.7), labels = c('A', 'B'), hjust = -1, label_size = 9, align = "v") +
  theme(plot.background = element_rect(color = "black", linewidth = 1),
        panel.background = element_rect(color = "black", linewidth = 1)) +
  annotation_custom(male, xmin = 0.18, xmax = 0.25, ymin = 0.31) +
  annotation_custom(female, xmin = 0.18, xmax = 0.24, ymin = 0.83)
TL_sex_SST_figcombined
```


# Save
```{r}
ggsave(TL_sex_SST_figcombined, file = "output/TL_sex_SST_figcombined.pdf", width = 80, height = 100, device = cairo_pdf, units = "mm", dpi = 300)
```


# Actual results


# Contrast - depth use by body size
```{r}
mod2 %>% 
emmeans(~Sex|TL, at = mod2_grid) %>% 
  regrid()

mod2 %>% 
emmeans(~Sex|TL, at = mod2_grid) %>% 
  regrid()%>% 
  pairs() %>% 
  confint()

mod2 %>% 
emmeans(~TL|Sex, at = mod2_grid) %>% 
  regrid() %>% 
  pairs() %>% 
  confint()
```

# Trend

```{r}
#sex_trends <- 
  mod2 %>% 
  emtrends(specs = pairwise~"Sex", var = "SST", infer = TRUE) %>% 
  regrid()

sex_trendsum <- summary(sex_trends, infer = TRUE)

sex_trendsum
```

```{r}
library(modelbased)
mod2 %>% estimate_slopes(trend = "SST", at = "Sex")
mod1 %>% ggeffect()
```


```{r}
# END #
```


