---
title: "Individual movement results - Sex * Size"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	warning = FALSE)
```


```{r include=FALSE}
#rm(list=ls()) # Clear out environment if necessary
source("code/packages.R")
```


This document digs down into the interactive effect of SST and Sex on the magnitude of daily depth movements in individual sharks

# Load
```{r}
load(file = "data/Rdata/shark_individual_daily_td.Rdata")
```

# Load model
```{r}
load(file = "data/Rdata/final_model.Rdata")
```



#  Mean TL 
```{r}
shark_individual_daily_td %>% 
  ungroup() %>% 
  group_by(Sex) %>% 
  summarise_at(vars(TL),
               list(min=min, 
                    mean = mean,
                    max=max))
```

# Overall trends:
```{r}
sst_maxd_final %>% 
  summary()

sex_trends <- sst_maxd_final %>% 
  emtrends(specs = pairwise~"Sex", var = "TL")

sex_trendsum <- summary(sex_trends, infer = TRUE)

sex_trendsum
```
Females respond with greater use of depth with increased body size
Males no response either way (probably due to small sample size)

# 1. Using intervals based on body size range in cohort

## Find what the range is
```{r}
shark_individual_daily_td %>% 
  dplyr::select(Sex, TL) %>% 
  unique() %>% 
  group_by(Sex) %>% 
  summarise(TL_max = max(TL),
            TL_min = min(TL))
```


```{r  fig.width = 160/25.4, fig.height = 100/25.4}
Sex_TL_grid <- shark_individual_daily_td %>% 
    with(list(Sex = levels(Sex),
              TL = c(125, 150, 175)))
  
newdata_Sex_TL <- sst_maxd_final %>% 
  emmeans(~TL|Sex, at = Sex_TL_grid) %>% 
  regrid() %>% 
  as_tibble() %>% 
  mutate(TL = factor(TL),
         Sex_TL = factor(paste(Sex, TL)))

sst_maxd_final
```


```{r  fig.width = 100/25.4, fig.height = 60/25.4}
(sex_TL_plot <- 
newdata_Sex_TL %>% 
  ggplot() +
  geom_linerange(aes(x = TL, y = response, ymin=asymp.LCL, ymax=asymp.UCL, group = Sex), colour = "grey40", position=position_dodge(0.5)) +
  geom_point(aes(x = TL, y = response, colour = Sex, fill = Sex, shape = Sex), colour = "grey20", position=position_dodge(0.5), size = 4) +
  scale_shape_manual(values = c(21,22)) +
  scale_fill_manual(values = c("darkblue", "dodgerblue")) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 140), breaks = c(0, 25, 50, 75, 100, 125)) +
  #theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  #ylab("Daily depth movement (m)\n(e.m. mean ± 95%CI)") + 
  xlab("\nTotal length (cm)") +
  theme(legend.position='none',
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.title = element_text(size = 8)))

```


# Save
```{r}
ggsave(sex_TL_plot, file = "output/Sex_TL_plot.pdf", width = 65, height = 50, units = "mm", dpi = 300)
```


# Combine with the Sex*SST plot
```{r  fig.width = 170/25.4, fig.height = 70/25.4}
SST_sex_TL_fig <- cowplot::plot_grid(Sex_SST_plot2, sex_TL_plot, nrow = 1, labels = c('A', 'B'), label_size = 9, align = "h", rel_widths = c(1.1, 0.6)) +
  theme(plot.background = element_rect(color = "black"))

SST_sex_TL_fig
```


# Save
```{r}
ggsave(SST_sex_TL_fig, file = "output/SST_sex_TL_fig.pdf", width = 170, height = 70, device = cairo_pdf, units = "mm", dpi = 300)
```





# 2. Plot with body size as continuous on X axis

```{r  fig.width = 160/25.4, fig.height = 100/25.4}
Sex_TL_grid <- shark_individual_daily_td %>% 
    with(list(Sex = levels(Sex),
              TL = seq_range(TL, n = 100)))
```


```{r  fig.width = 160/25.4, fig.height = 100/25.4}
newdata_Sex_TL <- sst_maxd_final %>% 
  emmeans(~TL|Sex, at = Sex_TL_grid) %>% 
  regrid() %>% 
  as_tibble() 
```


```{r  fig.width = 140/25.4, fig.height = 70/25.4}
#sex_TL_plot_cont <- 
newdata_Sex_TL %>% 
  ggplot() +
  geom_point(data = shark_individual_daily_td, aes(x = TL, y = daily_mean_depth, fill = Sex, colour = Sex), alpha = 0.05, position = position_jitter(width = 3)) +
  geom_line(aes(x = TL, y = response, colour = Sex, group = Sex), size = 1) +
  geom_ribbon(aes(x = TL, y = response, ymin=asymp.LCL, ymax=asymp.UCL, fill = Sex, group = Sex), alpha = 0.4) +
  scale_fill_manual(values = c("darkblue", "dodgerblue")) +
  scale_colour_manual(values = c("darkblue", "dodgerblue")) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 130), breaks = c(0, 25, 50, 75, 100, 125)) +
  #theme(plot.background = element_rect(fill = "transparent", colour = "black", linewidth = 1)) +
  ylab("Daily depth movement (m)\n(e.m. mean ± 95%CI)") + 
  xlab("\nTotal length (cm)") +
  theme(legend.position='none',
        legend.background = element_rect(fill = "white", colour = "black"),
        axis.title.x = element_text(size = 8),
        axis.text.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 8)) +
  facet_wrap(~ Sex)

sex_TL_plot_cont
```

# Combine with the Sex*SST plot
```{r  fig.width = 170/25.4, fig.height = 70/25.4}
SST_sex_TL_fig <- cowplot::plot_grid(Sex_SST_plot2, sex_TL_plot_cont, nrow = 1, labels = c('A', 'B'), label_size = 9, align = "h", rel_widths = c(1,1)) +
  theme(plot.background = element_rect(color = "black"))

SST_sex_TL_fig
```

# Save
```{r}
ggsave(SST_sex_TL_fig, file = "output/SST_sex_TL_fig2.pdf", width = 170, height = 70, device = cairo_pdf, units = "mm", dpi = 300)
```



```{r}
# END #
```